### 2024-10-07 kw/7-conclusion.tex

% !TEX root = main.tex
\chapter{Conclusion}
\label{ch:conclusion}

\noindent
In this work, we propose the first executable formal semantics for OpenQASM
2.0, a representative quantum circuit description language.
%
To address the nondeterministic nature of quantum circuits, our semantics
computes the probability distribution of possible outcomes.
%
We adopt MWI from the physics literature to compute the probability
distribution even for dynamic quantum circuits and optimize the semantics using
mixed states.
%
Additionally, we prove the physical consistency of the semantics, which implies
adherence to the postulates of quantum mechanics.
%
To simplify the proof, we define OpenQASMCore as a core language of OpenQASM
and desugar OpenQASM to OpenQASMCore.
%
Our evaluation demonstrates the effectiveness of the optimizations in improving
performance and the utility of the semantics as a testing oracle, finding five
previously unknown bugs in two real-world quantum computing platforms.