### 2024-10-07 kw/main.tex

% -*- TeX:UTF-8 -*-
%%
%% KAIST 학위논문양식 LaTeX용 (ver 0.5) 예시
%%
%% @version 0.4
%% @author  채승병 Chae,Seungbyung (mailto:chess@kaist.ac.kr)
%% @date    2004. 11. 12.
%%
%% @requirement
%% teTeX, fpTeX, teTeX 등의 LaTeX2e 배포판
%% + 은광희 님의 HLaTeX 0.991 이상 버젼 또는 홍석호 님의 HPACK 1.0
%% : 설치에 대한 자세한 정보는 http://www.ktug.or.kr을 참조바랍니다.
%%
%% @note
%% 기존에 널리 쓰여오던 차재춘 님의 학위논문양식 클래스 파일의 형식을
%% 따르지 않고 전면적으로 다시 작성하였습니다. 논문 정보 입력부분에서
%% 과거 양식과 다른 부분이 많으니 아래 예시에 맞춰 바꿔주십시오.
%%
%%
%% @acknowledgement
%% 본 예시 논문은 물리학과 박사과정 김용현 님의 호의로 제공되었습니다.
%%
%% -------------------------------------------------------------------
%% @information
%% 이 예제 파일은 hangul-ucs를 사용합니다. UTF-8 입력 인코딩으로
%% 작성되었습니다. hlatex의 hfont는 이용하지 않습니다. --2006/02/11
%% 본 템플릿은 전산학부 김민혁 교수에의해서 버그 수정되었습니다. -- 2016/11/25
%% 본 템플릿은 전산학부 김민혁 교수에의해서 추가 버그 수정되었습니다. -- 2023/06/15

% @class kaist.cls
% @options [default: doctor, korean, final]
% - doctor: 박사과정 | master : 석사과정
% - korean: 한글논문 | english: 영문논문
% - final : 최종판   | draft  : 시험판
% - pdfdoc : 선택하지 않으면 북마크와 colorlink를 만들지 않습니다.

\documentclass[master,english,final]{kaist-ucs} % 석사과정
% \documentclass[doctor,english,final]{kaist-ucs} % 박사과정

% If you want make pdf document (include bookmark, colorlink)
%\documentclass[doctor,english,final,pdfdoc]{kaist-ucs}

% kaist.cls 에서는 기본으로 dhucs, ifpdf, graphicx 패키지가 로드됩니다.
% 추가로 필요한 패키지가 있다면 주석을 풀고 적어넣으십시오,
%\usepackage{...}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[english]{babel}
\usepackage[numbers]{natbib}
% \usepackage{indentfirst}
\usepackage{booktabs}

\usepackage{mathpartir}
\usepackage[capitalise,nameinlink]{cleveref}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{mdframed}
\usepackage{pifont}
\usepackage{xfrac}
\usepackage{enumitem}

\usetikzlibrary{quantikz2}

\crefformat{chapter}{#2\S{}#1#3}
\Crefformat{chapter}{#2Chapter~#1#3}
\crefname{postulate}{Postulate}{Postulates}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\newcommand{\SUCC}{\ding{51}}
\newcommand{\FAIL}{\ding{55}}
\newcommand{\C}[1]{{\tt\small#1}}
\newcommand{\inred}{\color{red}}
\newcommand{\todo}{{\inred TODO}}
\newcommand{\embox}[1]{\mbox{\emph{#1}}}
\newcommand{\rep}[1]{\overline{#1}}
\newcommand{\ugate}[3]{\mathsf{U}(#1,#2,#3)}
\newcommand{\cxgate}{\mathsf{CX}}
\newcommand{\inop}{\mathsf{nop}}
\newcommand{\irotate}[4]{\ugate{#1}{#2}{#3}\ #4}
\newcommand{\icnot}[2]{\cxgate\ #1\ #2}
\newcommand{\imeasure}[2]{#2:=\mathsf{Measure}\ #1}
\newcommand{\iseq}[2]{#1;#2}
\newcommand{\iif}[3]{\mathsf{if}\ #1\rightsquigarrow#2\ \mathsf{then}\ #3}
\newcommand{\eval}[3]{#1\vdash#2\Rightarrow#3}
\newcommand{\sunitary}[2]{\embox{gate}(#1,#2)}
\newcommand{\smeasure}[2]{\embox{msr}(#1,#2)}
\newcommand{\sprob}[2]{\embox{prob}(#1,#2)}
\newcommand{\trace}[1]{\textrm{tr}(#1)}
\newcommand{\btrace}[1]{\textrm{tr}\big(#1\big)}
\newcommand{\qubits}[1]{\embox{qbts}(#1)}
\newcommand{\generalize}[2]{\embox{gen}_{#1}(#2)}
\newcommand{\mtrx}[1]{\begin{pmatrix}#1\end{pmatrix}}
\newcommand{\smtrx}[1]{\big(\begin{smallmatrix}#1\end{smallmatrix}\big)}
\newcommand{\Smtrx}[1]{\Big(\begin{smallmatrix}#1\end{smallmatrix}\Big)}
\newcommand{\identity}[1]{\embox{I}_{#1}}
\newcommand{\mrotate}[1]{\embox{U}_{#1}}
\newcommand{\mcnot}[1]{\embox{CX}_{#1}}
\newcommand{\mswap}[1]{\embox{Swap}_{#1}}
\newcommand{\mproj}[1]{\embox{Proj}_{#1}}
\newcommand{\dcreg}[2]{\mathsf{Creg}\ #1[#2]}
\newcommand{\dqreg}[2]{\mathsf{Qreg}\ #1[#2]}
\newcommand{\dgate}[4]{\mathsf{Gate}\ #1(#2)\ #3=#4}
\newcommand{\icustom}[3]{#1(#2)\ #3}
\newcommand{\ireset}[1]{\mathsf{Reset}\ #1}
\newcommand{\dom}[1]{\textrm{dom}(#1)}

% @command title 논문 제목(title of thesis)
% @options [default: (none)]
% - korean: 한글제목(korean title) | english: 영문제목(english title)
\title[korean] {양자 회로 시뮬레이터의 테스트 오라클로서의 물리적으로 일관된 실행 가능한 형식 의미론}
\title[english]{Physically-Consistent Executable Formal Semantics as a Testing Oracle for Quantum Circuit Simulators}

% @note 표지에 출력되는 제목을 강제로 줄바꿈하려면 \linebreak 을 삽입.
%       \\ 나 \newline 등을 사용하면 안됩니다. (아래는 예시)
%
%\title[korean]{탄소 나노튜브의 물리적 특성에 대한\linebreak 이론 연구}
%\title[english]{Theoretical study on physical properties of\linebreak
%                carbon nanotubes}
%
% If you want to begin a new line in cover, use \linebreak .
% See examples above.
%

% @command author 저자 이름
% @param   family_name, given_name 성, 이름을 구분해서 입력
% @options [default: (none)]
% - korean: 한글이름 | chinese: 한문이름 | english: 영문이름
% 한문 이름이 없다면 빈 칸으로 두셔도 됩니다.
%
%
% If you are a foreigner , write your name in korean or your korean name.
% If you can't write native character, you can make the chinese blank empty
% Write as follow
% \author[korean]{family name in korean}{given name in korean}
% \author[chinese]{family name in your native language}{given name in your native language}
% \author[english]{family name in english}{given name in english}
%
\author[korean] {이}{강 욱}
\author[korean2] {이}{강욱}    %이름을 붙여 써 주시기 바랍니다.
\author[chinese]{李}{康 旭}
\author[english]{Lee}{Kanguk}

% @command advisor 지도교수 이름 (복수가능)
% @usage   \advisor[options]{...한글이름...}{...영문이름...}{signed|nosign}
% @options [default: major]
% - major: 주 지도교수  | coopr: 공동 지도교수
\advisor[major]{류 석 영}{Sukyoung Ryu}{signed}
\advisor[major2]{류석영}{Sukyoung Ryu}{signed}    %한글 성과 한글 이름을 모두 붙여 써 주시기 바랍니다.

% [주의] 전산학부의 경우, 전공이름(Computer Science)을 적어주시기 바랍니다. 조직명(School of Computing) 적지 말아주세요!
\advisorinfo{Professor of Computer Science} %제출승인서에 들어가는 교수님 정보, advisor's information

%\advisor[coopr]{홍 길 동}{Gil-Dong Hong}{nosign}
%\advisor[coopr2]{홍길동}{Gil-Dong Hong}{nosign}    %한글 성과 한글 이름을 모두 붙여 써 주시기 바랍니다.
%
% 지도교수 한글이름은 입력하지 않아도 됩니다.
% You may not input advisor's korean name
% like this \advisor[major]{}{Chang, Kee Joo}{signed}
%

% @command department {학과이름}{학위종류} - 아래 규칙에 따라 코드를 입력
% @command department {department code}{degree field}
%
% department code
% 2. 석박사학위논문 작성 및 제출요령 4쪽 ~ 5쪽 참고
% 또는 kaist-ucs.cls 의 % @command department 참고

% science: 이학 | engineering: 공학 | business : 경영학
% 박사논문의 경우는 학위종류를 입력하지 않아도 됩니다.
% If you write Ph.D. dissertation, you cannot input degree field.
% The third parameter : a | b | c
% a: 소속된 학과만 쓰는 옵션 (학과에만 소속되어 있는 경우에는 무조건 a를 선택해야 함)
% b: 학과 아래의, 프로그램이나 학제전공에 소속되어 있을 경우에 학과와 프로그램을 함께 쓰는 옵션
% c: 학과 아래의, 프로그램이나 학제전공에 소속되어 있을 경우에 학과를 쓰지 않고 프로그램이나 학제전공의 이름만 쓰는 옵션
%
% a: it represents only the name of department. (if you aren't in the program under the department, must choose a)
% b: it represents the names of department and the program that is under the department (consider this when you are in the program not only department)
% c: it represents only the name of program that is under the department (consider this when you are in the program not only department)
\department{CS}{engineering}{a}

% @command referee 심사위원 (석사과정 3인, 박사과정 5인)
\referee[1]{류 석 영}
\referee[2]{유 신}
\referee[3]{강 지 훈}
% \referee[4]{가 동 호}
% \referee[5]{박 태 현}
% \referee[5] {Barack Obama}
% Of course english name is available

% @command approvaldate 지도교수논문승인일
% @param   year,month,day 연,월,일 순으로 입력
\approvaldate{2023}{12}{12}

% @command refereedate 심사위원논문심사일
% @param   year,month,day 연,월,일 순으로 입력
\refereedate{2023}{12}{12}

% @command gradyear 졸업년도
\gradyear{2024}

% 본문 시작
\begin{document}

% \theoremstyle{acmdefinition}
\newtheorem{postulate}[theorem]{Postulate}

% 앞표지, 속표지, 학위논문 제출승인서, 학위논문 심사완료 검인서는
% 클래스 옵션을 final로 지정해주면 자동으로 생성되며,
% 반대로 옵션을 draft로 지정해주면 생성되지 않습니다.

% 논문 서지, 초록, 핵심 낱말, 영문 초록, 영어 핵심 낱말 (Information of thesis, abstract in korean, keywords in korean, abstract in english, keywords in english)
\thesisinfo
%% Letters of abstract in korean must be less than 500 and words of abstract in english must be less than 300.
%% Number of keywords must be less than 6.
%% Don't write english letters in the abstract in korean.
\begin{summary}
	%
	현재 양자 프로그램 개발에서는 개발자들이 양자 회로 시뮬레이터에 많이 의존하고 있다.
	%
	따라서 시뮬레이터의 정확성을 보장하는 것은 매우 중요하다.
	%
	그러나 양자 회로를 기술하는 언어에 대한 형식적 의미론이 없기 때문에 시뮬레이터의 정확성에 대한 추론이 어렵다.
	%
	본 논문에서는 대표적인 양자 회로 기술 언어인 OpenQASM 2.0에 대한 실행 가능한 형식 의미론을 제시한다.
	%
	양자 회로의 독특한 특성으로 인해, 형식 의미론을 정의할 때 크게 두 가지 어려움이 있다.
	%
	첫째, 시뮬레이터의 비결정성에도 불구하고 실행 가능한 형식 의미론은 결정론적인 결과를 내야 한다.
	%
	이를 위해 물리학계에서 제시된 다세계 해석을 이용하여 결과의 확률 분포를 계산하는 형식 의미론을 제시한다.
	%
	둘째, 실행 가능한 형식 의미론은 물리적 일관성을 보장해야 한다.
	%
	이를 위해 OpenQASM의 핵심 언어인 OpenQASMCore를 제시하고, OpenQASM에서 OpenQASMCore로의 변환(디슈가링)을 정의한다.
	%
	형식 의미론과 물리적 일관성에 대한 증명은 Coq에서 정형 검증되었다.
	%
	본 논문에서는 실행 가능한 형식 의미론의 평가로서 두 가지 최적화 기법의 효과적인 성능 향상과, 실제 양자 프로그래밍 플랫폼에서 5개의
	새로운 버그를 찾아내는 등의 실행 가능한 형식 의미론의 테스트 오라클로서의 유용성을 보여준다.
	%
\end{summary}

\begin{Korkeyword}
	양자 회로, 양자 프로그래밍 언어, 형식 의미론, 테스트 오라클, OpenQASM, OpenQASMCore, 다세계 해석, 양자 시뮬레이터,
	물리적 일관성
\end{Korkeyword}

\begin{abstract}
	In the current practice of quantum program development, programmers heavily
	rely on quantum circuit simulators.
	%
	Ensuring the correctness of these simulators is therefore crucial.
	%
	However, the lack of formal semantics for quantum circuit description
	languages, which are used to describe quantum circuits, hinders reasoning about
	simulator correctness.
	%
	In this work, we present the first executable formal semantics, which can
	function as a testing oracle, for OpenQASM 2.0, a representative quantum
	circuit description language.
	%
	The unique characteristics of quantum circuits present challenges in defining
	formal semantics.
	%
	First, executable semantics must yield deterministic results despite the
	nondeterminism of simulators.
	%
	To achieve this, our semantics computes the probability distribution of
	outcomes using the \emph{many-worlds interpretation} (MWI) from the physics
	literature.
	%
	Additionally, we propose two optimizations for the semantics.
	%
	Second, ensuring \emph{physical consistency} is essential, meaning the
	semantics must adhere to physical laws.
	%
	To facilitate this, we introduce OpenQASMCore, a core language for OpenQASM,
	simplifying the proof process, and define desugaring from OpenQASM to
	OpenQASMCore.
	%
	The semantics and proofs are mechanized in Coq.
	%
	Our evaluation shows the optimizations' effectiveness in improving performance
	and the utility of our semantics as a testing oracle by identifying five
	previously unknown bugs in two real-world quantum programming platforms.
	%
\end{abstract}

\begin{Engkeyword}
	Quantum circuit, quantum programming language, formal semantics, testing
	oracle, OpenQASM, OpenQASMCore, many-worlds interpretation, Quantum simulator, and
	physical consistency.
\end{Engkeyword}

\addtocounter{pagemarker}{1}                 % 백색별지분을 고려
\newpage

% 목차 (Table of Contents) 생성
\tableofcontents

% 표목차 (List of Tables) 생성
\listoftables

% 그림목차 (List of Figures) 생성
\listoffigures

% 위의 세 종류의 목차는 한꺼번에 다음 명령으로 생성할 수도 있습니다.
%\makecontents

%% 이하의 본문은 LaTeX 표준 클래스 report 양식에 준하여 작성하시면 됩니다.
%% 하지만 part는 사용하지 못하도록 제거하였으므로, chapter가 문서 내의
%% 최상위 분류 단위가 됩니다.
%% You cannot use 'part'

\include{1-intro}

\include{2-openqasmcore}

\include{3-consistency}

\include{4-openqasm}

\include{5-evaluation}

\include{6-related}

\include{7-conclusion}

\chapter[Method]{Method}
\label{ch:ourmethod}
\section{Quantum Communication}
To enhance the performance gain of original cognitive radio networks (CRNs),
leveraging cooperative diversity has attracted a lot of attention
\cite{SOCA1}...

\section{Examples}

\section{Applications of Quantum Communication}
\label{ch:application}
\noindent
Despite its potential to improve the throughput, spatial domain diversity was not fully considered in the studies of original CCRNs. Utilizing the spatial domain for the communications, the concept of MIMO has been adopted in many cases to increase the wireless capacity \cite{EF1,FD2}...

Let us consider an MIMO-CCRN with one PL and one SL. Each PU has one legacy
antenna and each SU has two STAR antennas. The duration of one communication
time frame is divided into two phases \cite{RVP2,ML2} ...
%%
%% 표 삽입 예시
%% Example. how to insert table
%%
\begin{table}[t]
	\caption[Enter the caption title here]{Energy stability $E$ (eV) per molecule of all meta-stable
		isomer states of C$_{60}$ opening process for forming the (5,5) cap.
		In the SW-I and SW-II, both ferromagnetic (Ferro) and paramagnetic (Para)
		spin configurations are obtained, whereas only non-magnetic configuration
		is obtained in the BF, SW-III, and CAP(5,5).
		$M$ is total magnetization $n_{\rm up}$-$n_{\rm down}$ in unit of $\mu_B$, where
		$n_{\rm up(down)}$ is the number of up (down) spins.
	}
	\label{mag-tab1}
	\begin{center}
		\begin{tabular} {ccccccccccc}
			\hline\hline
			 &               & BF & \multicolumn{2}{c}{SW-I} &       & \multicolumn{2}{c}{SW-II} & SW-III & CAP    &               \\
			\cline{4-5} \cline{7-8}
			 &               &    & Para                     & Ferro &                           & Para   & Ferro  &      &      & \\
			\hline
			 & $E$ (eV)      & 0  & 7.796                    & 7.832 &                           & 10.418 & 10.408 & 11.5 & 13.2 & \\
			 & $M$ ($\mu_B$) & 0  & 0                        & 1.94  &                           & 0      & 2.06   & 0    & 0    & \\
			\hline\hline
		\end{tabular}
	\end{center}
\end{table}

%%
%% 그림 삽입 예시
%% Example. how to insert graph
%%
%% Note. 가급적 \includegraphics 명령을 사용하십시오.
%% Recommen : Use \includegraphics to insert graph.
%%
\begin{figure}[t]
	\centerline{\includegraphics[width=12.5cm]{sample-fig1}}
	\caption[Enter the caption title here]{ Ball-and-stick models of meta-stable isomers in
		cage opening process from a C$_{60}$ buckminsterfullerene
		to a (5,5) capsule. We name them BF and CAP(5,5).
		Depending on the number of the Stone-Wales (SW) transformation,
		we call the intermediate isomers with SW-I, SW-II, and SW-III.
		Highlighted atoms are undercoordinated except BF.
	} \label{mag-fig1}
\end{figure}

\section{Discussion and Future Work}
\noindent
The achievable rates can be calculated by finding the statistics of the five signals transmitted that maximize the mutual information between $s_{t,XY}$ and $y_{t,XY}$ for $(X,Y)=(T,C), (C,N)$, and $(N,C)$ when $t=1$, and $(X,Y)=(C,R), (C,N)$, and $(N,C)$ when $t=2$ \cite{SOCA2,EF2}...

\chapter{Conclusion}
\noindent
We have proposed a novel FD MIMO-CCRN framework providing a reasonable performance improvement compared with the conventional MIMO-CCRN framework...
%%
%% 참고문헌 시작
%% bibliography
%% It can be changed but should include sufficient information.
% \bibliographystyle{abbrvnat}
\bibliographystyle{plainnat}
\bibliography{references}

%%
%% 감사의 글 시작
%% Acknowledgement
%%
% @command acknowledgement 감사의글
% @options [1 | 2 | 3 |4 ]
% - 1 : 본문과 감사의 글이 둘 다 한글일 때  | 2 : 본문은 한글인데 감사의 글이 영어일 때 | 3 :  본문과 감사의 글이 둘 다 영어일 때  | 4 : 본문은 영어인데 감사의 글이 % 한글일 때
%% It is optional.

\acknowledgment[4]
언제나 저를 바른 길로 이끌어 주시는 송익호 교수님께 큰 고마움을 느낍니다.
끝으로 오늘의 제가 있을 수 있도록 사랑으로 키워 주신 가족들에게 감사드립니다.
저의 이 작은 결실이 그분들께 조금이나마 보답이 되기를 바랍니다.

%%
%% 약력 시작
%% Curriculum Vitae
%%
% @command curriculumvitae 이력서
% @options [1 | 2 | 3 |4 ]
% - 1 : 본문과 약력이 둘 다 한글일 때  | 2 : 본문은 한글인데 약력이 영어일 때 | 3 :  본문과 약력이 둘 다 영어일 때  | 4 : 본문은 영어인데 약력이 한글일 때
%% It is optional and you can change form of this in the class file if you want.
\curriculumvitae[4]

% @environment personaldata 개인정보
% @command     name         이름
%              dateofbirth  생년월일
%              birthplace   출생지
%              domicile     본적지
%              address      주소지
%              email        E-mail 주소
% - 위 6개의 기본 필드 중에 이력서에 적고 싶은 정보를 입력
% input data only you want
\begin{personaldata}
	\name       {이 강 욱}
	\dateofbirth{1999}{08}{01}
	\birthplace {...}
	\address    {...}
\end{personaldata}

% @environment education 학력
% @options [default: (none)] - 수학기간을 입력
\begin{education}
	\item[2015. 3.\ --\ 2017. 2.] 고등학교 (2년 수료)
	\item[2017. 3.\ --\ 2022. 8.] 한국과학기술원 물리학과 (학사)
	\item[2022. 9.\ --\ 2024. 2.] 한국과학기술원 전산학부 (석사)
\end{education}

% @environment career 경력
% @options [default: (none)] - 해당기간을 입력
\begin{career}
	\item[2021. 3.\ --\ 2021. 8.] 쓰레기통 청소
\end{career}

% @environment activity 학회활동
% @options [default: (none)] - 활동내용을 입력
%%    \begin{activity}
%%        \item J. Choi, \textbf{Yong-Hyun Kim}, K.J. Chang, and D. Tomanek,
%%             \textit{Occurrence of itinerant ferromagnetism in C/BN superlattice
%%             nanotubes}, 5th Asian Workshop on First-Principles Electronic
%%             Structure Calculations, Seoul (Korea), October., 2002.
%%    \end{activity}

% @environment publication 연구업적
% @options [default: (none)] - 출판내용을 입력
\begin{publication}
	%\item \textbf{Yong-Hyun Kim}, J. Choi, K.J. Chang, and D. Tomanek,
	%     \textit{Magnetic instability in partly opened C$_{60}$ isomers},
	%     in preparation.
	\item H.-K. Min, Y. Hou, {\bf S. Park}, and I. Song, ``A computationally efficient
	scheme for feature extraction with kernel discriminant analysis," \textit{Patt.
		Recogn.}, vol.~50, no.~2, pp.~45-55, Feb. 2016 (to be published).
\end{publication}

\label{paperlastpagelabel}     % <-- 추가 부분: 마지막 페이지 위치 지정
%% 본문 끝
\end{document}