### 2024-10-07 paper/1-intro.tex
Write main.tex

% !TEX root = main.tex

\chapter{Introduction}
\label{ch:intro}
\noindent