### 2024-10-15 paper/5-eval.tex
WIP: Formalize AL semantics

% !TEX root = main.tex

\chapter{Evaluation}
\label{ch:eval}
\noindent

* sem
  - pass all official test includes wasm 3.0
  - pass exit semantics testing
  - faster?
* type syste
  - structure aware rule -> type based rule
  - byte O