### 2024-10-29 paper/5-eval.tex
Add space

% !TEX root = main.tex

\chapter{Evaluation}
\label{ch:eval}
\noindent

% an example of buggy test case
\textbf{Example 5.1}
\begin{verbatim}
  // function definition of $br-returning
  (func  $brAndReturning (br 0) (unreachable))

  // function call
  (call $brAndReturning)
\end{verbatim}

Example 5.1 is the minimul example that produces a wrong result in \spectecp{}.
It contains a function that has only a \texttt{br} instruction.
According to the \officialp{}, when the function \texttt{\$brAndReturning} is
called, a frame is pushed to the stack, and the block of the function
body is entered with a label.
When the \texttt{br} is executed, it pops the label from the stack.
As the end of the function is reached after the execution, returning from a
function is performed.
As a result, the frame is popped, so the call actually does nothing.

% explaination of the bug in SpecTec prose
\begin{align}
  &c^* \vdash call(\$brAndReturning) \label{eq:bug-1} \\
&\leadsto
  Label(\epsilon), ~ Frame, ~ c^* \vdash (br ~ 0) ~ (unreachable), ~ end \label{eq:bug-2} \\
&\leadsto
  Frame, ~ c^* \vdash \epsilon \label{eq:bug-3}
\end{align}

In \spectecp{} the call instruction in \cref{eq:bug-1} pushes a frame and
enters the function body \cref{eq:bug-2}.
When the \texttt{br} is executed, it pops the label and removes the input
instructions until the end in \cref{eq:bug-3}.
As a result, the frame is remained after the executing the call.

% simplified version of AL algorithm of call, br, and returning from a function
\begin{verbatim}
Call fname:
  PushCtx Frame
  Enter get_function_body(fname) with Label(\epsilon)

Br 0:
  PopCtx Label(instr*)
  ExecuteSeq instr*

Returning from a function:
  PopCtx Frame
\end{verbatim}

\red{TODO: Change to array} \\
\red{TODO: Change naming for $c$}
\begin{align}
  &c^* \vdash call(\$brAndReturning); ~ \wasm ~ (n, ~ c)
  \label{eq:resolve-1} \\
  &\leadsto c^* \vdash \epsilon; ~ \al ~ ([ ~ \pushctxi, ~ \enteri ~ ], ~ 0, ~ \wasm ~ (n, ~ c))
  \label{eq:resolve-2} \\
  &\leadsto Frame, ~ c^* \vdash \epsilon; ~ \al ~ ([ ~ \enteri ~ ], ~ 1, ~ \wasm ~ (n, ~ c))
  \label{eq:resolve-3} \\
  &\leadsto Label(\epsilon), ~ Frame, ~ c^* \vdash (br ~ 0), ~ (unreachable), ~ (end ~ 2); ~
    \wasm ~ (2, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c)))
  \label{eq:resolve-4} \\
  &\leadsto Label(\epsilon), ~ Frame, ~ c^* \vdash (unreachable), ~ (end ~ 2); \\
  &\quad\quad \al ~ ([ ~ \red{\symbol{PopCtxI}}, ~ \executeseqi ~ ], ~ 0, ~ \wasm ~ (2, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c))))
  \label{eq:resolve-5} \\
  &\leadsto Frame, ~ c^* \vdash (end ~ 1); ~
    \al ~ ([ ~ \executeseqi ~ ], ~ 0, ~ \wasm ~ (1, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c))))
  \label{eq:resolve-6} \\
  &\leadsto Frame, ~ c^* \vdash (end ~ 1); ~
    \al ~ (\epsilon, ~ 0, ~ \wasm ~ (1, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c))))
  \label{eq:resolve-7} \\
  &\leadsto Frame, ~ c^* \vdash (end ~ 1); ~ \wasm ~ (1, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c)))
  \label{eq:resolve-8} \\
  &\leadsto Frame, ~ c^* \vdash \epsilon; ~
    \al ~ ([ ~ \red{\symbol{PopCtx}} ~ ], ~ 0, ~ \wasm ~ (0, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c))))
  \label{eq:resolve-9} \\
  &\leadsto c^* \vdash \epsilon; ~
    \al ~ (\epsilon, ~ 0, ~ \wasm ~ (0, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c))))
  \label{eq:resolve-10} \\
  &\leadsto c^* \vdash \epsilon; ~ \wasm ~ (0, ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c)))
  \label{eq:resolve-11} \\
  &\leadsto c^* \vdash \epsilon; ~ \al ~ (\epsilon, ~ 0, ~ \wasm ~ (n, ~ c))
  \label{eq:resolve-12} \\
  &\leadsto c^* \vdash \epsilon; ~ \wasm ~ (n, ~ c)
  \label{eq:resolve-13}
\end{align}