### 2024-11-16 paper/1-intro.tex
WIP: write intro

% !TEX root = main.tex

\chapter{Introduction}
\label{ch:intro}
\noindent

\begin{itemize}
\item Wasm
  \begin{itemize}
  \item is low-level bytecode language, platform-indep, ...
  \item is compile target for Web applications
  \item is used in wide area such as IoT, ...
  \end{itemize}
\item Wasm Spec
  \begin{itemize}
  \item descibes syntax and semantics of Wasm
  \item should be rigorous because ... (brief explanation for two format)
  \item has two format to describe semantics (formal\&prose notation)
  \item formal notation is written in math expression: concise, clear, amenable to proofs, ...
  \item prose notation is step by step instruction written English: vague (in some part), amenable to developing engines
  \end{itemize}
\item SpecTec
  \begin{itemize}
  \item is proposed to alleviate spec writing process
  \item is accepted by Wasm committee to be an official Wasm specification tool
  \item define AL to describe Wasm semantics in prose notation
  \item AL is designed to be similar to prose notation (describe with fig)
  \item implement AL interpreter to make the spec executable
  \item executable spec passes official wasm test $\leftarrow$ correctness
  \end{itemize}
\item Problem
  \begin{itemize}
  \item no semantics in AL
  \item prose notation is still vague unless read AL interpreter code
  \item ???
  \end{itemize}
\item Contribution
  \begin{itemize}
  \item formalize AL syntax$\&$semantics
  \item make prose notation clear to understand
  \item ???
  \end{itemize}
\end{itemize}

% breif overview
WebAssembly (Wasm)~\cite{wasm} is a low-level bytecode language that is safe, fast,
portable, and compact.
It is widely used as a compilation target for web applications.
Beyond web development, Wasm's advantages are also deployed in areas such as
edge computing~\cite{wasm-edge}, IoT~\cite{wasm-iot}, and
blockchain~\cite{wasm-block}.

% risk of implementation divergence
There are numerous Wasm engines, with all major browsers implementing their own
multi-tiered versions~\cite{v8, spidermonkey, webkit}.
Additionally, specialized engines target specific domains, such as embedded
systems and edge computing.
However, ensuring portability across these diverse implementations introduces
the risk of divergence among them.

% rigorous standardization -> spec is rigorous: formal notation & prose notation
To address this challenge, Wasm has been rigorously standardized by the
W3C~\cite{wasm-w3c}.
The Wasm specification is particularly rigorous, describing its semantics in
two complementary forms: formal notation and prose notation.
The formal notation employs mathematical rules to succinctly define the
semantics, supporting proofs such as type soundness.
Conversely, the prose notation uses pseudocode-like algorithms to explain the
semantics through step-by-step instructions.
Since most Wasm users and engine developers are less familiar with mathematical
formalism, they primarily rely on the prose notation.

% challenging specification process
The demanding standardization process places a significant burden on
specification authors.
Crafting this specification document is labor-intensive, and as Wasm evolves,
the manual effort required becomes increasingly challenging to scale.
Moreover, the dual requirement of maintaining both the formal notation and
prose notation exacerbates these difficulties.
The formal notation is written in LaTeX, and the prose notation is authored
in reStructuredText; neither of which is particularly user-friendly for
collaborative review.
This lack of accessibility in the specification's tooling increases the
likelihood of inconsistensies and errors, further complicating the
standardization effort.

% SpecTec
To alleviate this problem, we had been developing SpecTec, a framework for
mechanizing WebAssembly specification.
It provides DSL to define Wasm syntax and semantics in a declarative style
similar to the formal notation.
It type-checks DSL to prevent meta-level specification errors, and generates
many artifacts including the specification document, a Wasm interpreter, and
mechanized definitions for theorem provers.

% AL
To express

% problem
However, AL cannot describe Wasm control flow correctly

% contribution
build AL's computation model for Wasm, formalize AL semantics