### 2024-11-20 paper/spectec.tex
Change file name

% !TEX root = main.tex

\chapter{SpecTec}
\label{ch:spectec}
\noindent

\newcommand{\seq}[1]{#1^*}

%% Syntax
\section{Syntax of AL}
\label{syntax}

{
\renewcommand{\arraystretch}{0.8835}  % Decrease row spacing locally
\begin{align*}
\begin{array}{lcccrlr}
%
% Algorithm
  \text{Algorithm}\quad& A &\ni& a &::=& ~ \rel ~ (s, \seq e, \seq i) ~ | ~ \fun ~ (s, \seq e, \seq i) \\
%
% Instruction
  \text{Instruction}\quad& I &\ni& i &::=& ~ \ifi ~ (e, \seq i, \seq i) &\quad\text{(If)} \\
    &&&& | & ~ \eitheri ~ (\seq i, \seq i) &\quad\text{(Either)} \\
    &&&& | & ~ \enteri ~ (e, e) &\quad\text{(Enter)} \\
    &&&& | & ~ \pushctxi ~ e &\quad\text{(Push Context)} \\
    &&&& | & ~ \pushi ~ e &\quad\text{(Push)} \\
    &&&& | & ~ \popctxi ~ e &\quad\text{(Pop Context)} \\
    &&&& | & ~ \popi ~ e &\quad\text{(Pop)} \\
    &&&& | & ~ \popni ~ (e, e) &\quad\text{(Pop N)} \\
    &&&& | & ~ \popalli ~ e &\quad\text{(Pop All)} \\
    &&&& | & ~ \leti ~ (e, e) &\quad\text{(Let)} \\
    &&&& | & ~ \trapi &\quad\text{(Trap)} \\
    &&&& | & ~ \returnreli &\quad\text{(Return Relation)} \\
    &&&& | & ~ \returnfuni ~ e &\quad\text{(Return Function)} \\
    &&&& | & ~ \executei ~ e &\quad\text{(Execute)} \\
    &&&& | & ~ \calli ~ (s, s, \seq{e}) &\quad\text{(Let Call)} \\
    &&&& | & ~ \replaceframei ~ (p^+, e) &\quad\text{(Replace Frame)} \\
    &&&& | & ~ \replacestorei ~ (p^+, e) &\quad\text{(Replace Store)} \\
%
% Expression
  \text{Expression}\quad& E &\ni& e &::=& ~ \vare ~ s &\quad\text{(Variable)} \\
    &&&& | & ~ \nume ~ n &\quad\text{(Number)} \\
    &&&& | & ~ \boole ~ b &\quad\text{(Boolean)} \\
    &&&& | & ~ \fnamee ~ s &\quad\text{(Function Name)} \\
    &&&& | & ~ \liste ~ \seq e &\quad\text{(List)} \\
    &&&& | & ~ \stre ~ \seq{(s, e)} &\quad\text{(Record)} \\
    &&&& | & ~ \tupe ~ \seq e &\quad\text{(Tuple)} \\
    &&&& | & ~ \casee ~ (s, \seq e) &\quad\text{(Tagged Tuple)} \\
    &&&& | & ~ \une ~ (unop, e) &\quad\text{(Unary Operation)} \\
    &&&& | & ~ \bine ~ (binop, e, e) &\quad\text{(Binary Operation)} \\
    &&&& | & ~ \acce ~ (e, p) &\quad\text{(Access)} \\
    &&&& | & ~ \upde ~ (e, p^+, e) &\quad\text{(Update)} \\
    &&&& | & ~ \cate ~ (e, e) &\quad\text{(Concatenation)} \\
    &&&& | & ~ \compe ~ (e, e) &\quad\text{(Composition)} \\
    &&&& | & ~ \meme ~ (e, e) &\quad\text{(Membership)} \\
    &&&& | & ~ \choosee ~ e &\quad\text{(Choose)} \\
    &&&& | & ~ \lene ~ e &\quad\text{(Length)} \\
    &&&& | & ~ \iscaseofe ~ (e, s) &\quad\text{(Check Tag)} \\
    &&&& | & ~ \getcurctxe  &\quad\text{(Get Current Context)} \\
    &&&& | & ~ \ctxkinde ~ s &\quad\text{(Context Kind)} \\
    &&&& | & ~ \itere ~ (e, iter, \seq{s}) &\quad\text{(Iteration)} \\
    &&&& | & ~ \matche ~ (e, e) &\quad\text{(Match)} \\
    &&&& | & ~ \hastypee ~ (e, s) &\quad\text{(Has Type)} \\
\end{array}
\end{align*}
}

\newpage
\begin{align*}
\begin{array}{lcccrlr}
%
% Path
  \text{Path}\quad& P &\ni& p &::=& ~ \idxp ~ e ~ | ~ \slicep ~ (e, e) ~ | ~ \dotp ~ s \\
%
% Iter
  \text{Iter}\quad& Iter &\ni& iter &::=& ~ \listiter ~ | ~ \listniter ~ e ~|~ \listidxiter ~ (s, e) \\
%
% Operator
  \text{Unary Operator}\quad& Unop &\ni& unop &::=& ~ \notop ~ | ~ \minusop \\
  \text{Binary Operator}\quad& Binop &\ni& binop &::=& ~ \addop ~ | ~ \subop ~ | ~ \mulop \\
    &&&& | & \divop ~ | ~ \modop ~ | ~ \expop \\
    &&&& | & \implop ~ | ~ \equivop ~ | ~ \andop \\
    &&&& | & \orop ~ | ~ \eqop ~ | ~ \neop ~ | ~ \ltop \\
    &&&& | & \gtop ~ | ~ \leop ~ | ~ \geop \\
% Primitives
  \text{String}\quad& \mathbb S &\ni& s \\
  \text{Integer}\quad& \mathbb Z &\ni& n \\
  \text{Boolean}\quad& \mathbb B &\ni& b \\
\end{array}
\end{align*}

% algorithm
An algorithm is either an AL relation or an AL function, corresponding to the
relation and the function definitions in the DSL.
% instruction
An AL instruction is \ifi{} for conditionally choosing a branch, \eitheri{} for
nondeterministically choosing a branch, \enteri{} for entering a block,
\pushctxi{} for pushing a Wasm context, \pushi{} for pushing a Wasm value,
\popctxi{} for popping a Wasm context, \popi{} for popping a Wasm value,
\popni{} for popping $n$ Wasm values, \popalli{} for popping all Wasm values
within a Wasm context, \leti{} for let binding, \trapi{} for trapping,
\returnreli{} for returning from an AL relation, \returnfuni{} for returning
from an AL function, \executei{} for executing given Wasm instructions,
\calli{} for an AL function call with let binding, \replaceframei{} for
replacing a frame, or \replacestorei{} for replacing a store.
% expression
An expression is a variable, number, boolean, AL function name, list, record
(or struct), tuple, tagged tuple, unary or binary operation, data structure
access or update, list or record concatenation, membership check, element
selection, length retrieval, tag check, Wasm context retrieval, Wasm context
kind check, iteration, match relation, or type relation.

\begin{align*}
\begin{array}{lcccrlr}
%
% State
  \text{State}\quad& \Sigma &\ni& \sigma & ::=& ~ \seq a, w, k \\
%
% Wasm state
  \text{Wasm State}\quad& W &\ni& w &::=& ~ \seq{we}, \seq{wi}, sto \\
%
% Wasm Entry
  \text{Wasm Entry}\quad& WE &\ni& we &::=& ~ wv ~ | ~ wc \\
%
% Wasm Value
  \text{Wasm Value}\quad& WV &\ni& wv &::=& ~ \casev ~ (s, \seq v) \\
%
% Wasm Context
  \text{Wasm Context}\quad& WC &\ni& wc &::=& ~ \casev ~ (\text{``Label"}, \seq v) ~ | ~ \casev ~ (\text{``Frame"}, \seq v) \\
%
% Wasm Instruction
  \text{Wasm Instruction}\quad& WI &\ni& wi &::=& ~ \casev ~ (s, \seq v) \\
%
% Store
  \text{Store}\quad& Store &\ni& sto &::=& ~ \seq{(s, v)} \\
%
% Continuation
  \text{Continuation}\quad & K &\ni& k &::=& ~ \mt &\quad\text{(Empty)} \\
    &&&& | & ~ \toplevelcall ~ (s, \seq v) &\quad\text{(Top-level Call)} \\
    &&&& | & ~ \call ~ (s, s, \seq v, k) &\quad\text{(Call)} \\
    &&&& | & ~ \exe ~ (\seq{wi}, k) &\quad\text{(Execute)} \\
    &&&& | & ~ \wasm ~ (\seq{wc}, k) &\quad\text{(Wasm)} \\
    &&&& | & ~ \algo ~ c &\quad\text{(Algorithm)} \\
    &&&& | & ~ \ret ~ (v, k) &\quad\text{(Return)} \\
%
% Context
  \text{Context}\quad& C &\ni& c &::=& ~ (\seq i, \mu, \seq{wc}, k) \\
%
% Environment
  \text{Environment}\quad& M &\ni& \mu &::=& ~ \seq{[s \mapsto v]} \\
\end{array}
\end{align*}

\newpage
\begin{align*}
\begin{array}{lcccrlr}
%
% Value
  \text{Value}\quad& V &\ni& v &::=& ~ \numv ~ n &\quad\text{(Number)} \\
    &&&& | & ~ \boolv ~ b &\quad\text{(Boolean)} \\
    &&&& | & ~ \fnamev ~ s &\quad\text{(Function Name)} \\
    &&&& | & ~ \listv ~ \seq v &\quad\text{(List)} \\
    &&&& | & ~ \strv ~ \seq{(s, v)} &\quad\text{(Record)} \\
    &&&& | & ~ \tupv ~ \seq v &\quad\text{(Tuple)} \\
    &&&& | & ~ \casev ~ (s, \seq v) &\quad\text{(Tagged Tuple)} \\
    &&&& | & ~ \trapv &\quad\text{(Trap)} \\
    &&&& | & ~ \storev &\quad\text{(Store)} \\
\end{array}
\end{align*}

% State
An AL state consists of a sequence of algorithms, a Wasm state, and a
continuation.
% Wasm State
A Wasm state consists of an interleaved stack, a Wasm instruction stack, and a
store.
% Wasm Entry
An interleaved stack contains two types of entries: a Wasm value and a Wasm
context.
% Wasm Value & Wasm Context & Wasm Instruction
Each of a Wasm value, a Wasm context, and a Wasm instruction is expressed as an
AL value $\casev$.
In particular, the tag of a Wasm context is either ``Label" or ``Frame".
% Store
A store is expressed as a sequence of pairs, each consisting of a field name
and an AL value.
% Continuation
A continuation is \mt{} for an empty continuation, \toplevelcall{} for
top-level function call, \call{} for a function call with let binding, \exe{}
for executing given Wasm instructions, \wasm{} for executing a Wasm instruction
in the instruction stack, \algo{} for interpreting AL instructions in an
algorithm, or \ret{} for returning from an AL function.
% Env
An environment is a finite mapping from variable names to AL values.
% Context
An AL context consists of a sequence of Wasm instruction, an environment, a
sequence of Wasm contexts, and a continuation.
% Value
An AL value is a number, boolean, function name, list, record, tuple, tagged tuple,
trap, or store.

% Terminology
In this section, AL-specific terms (state, context, value, function, and
instruction) are referred to simply as state, context, value, function, and
instruction, while Wasm terms remain unchanged.

\newpage
%% Semantics
\section{Semantics of AL}
\label{semantics}

%% Continuation

\begin{gather*}
\boxed{\leadsto \subseteq \Sigma \times \Sigma} \\
%
% TopLevelCall
\newline \\
  \\
  \hline
  (\seq{a}, w, \toplevelcall ~ (s, \seq v)) \leadsto (\seq{a}, w, \createalgo(\seq a, s, \seq v, \mt)) \\
%
% Call
\newline \\
  \\
  \hline
  (\seq{a}, w, \call ~ (s_{var}, s_{name}, \seq v, k)) \leadsto
  (\seq{a}, w, \createalgo(\seq a, s_{name}, \seq v, \call ~ (s_{var}, s_{name}, \seq v, k))) \\
%
% Execute-empty
\newline \\
  \\
  \hline
  (\seq{a}, w, \exe ~ (\epsilon, k)) \leadsto (\seq{a}, w, k) \\
%
% Execute-instr
\newline \\
  \casev ~ (s, \seq v) = wi \\
  \hline
  (\seq{a}, w, \exe ~ (wi ~ \seq{wi}, k)) \leadsto
  (\seq{a}, w, \createalgo(\seq a, s, \seq v, \exe ~ (\seq{wi}, k))) \\
%
% Wasm-empty
\newline \\
  \\
  \hline
  (\seq{a}, w, \wasm ~ (\epsilon, k)) \leadsto (\seq{a}, w, k) \\
%
% Wasm-instr
\newline \\
  (w', \casev ~ (s, \seq v)) = \popwasminstr(w) \\
  \hline
  (\seq{a}, w, \wasm ~ (wc^+, k))
  \leadsto
  (\seq{a}, w', \createalgo(\seq a, s, \seq v, \wasm ~ (wc^+, k))) \\
%
% Al-empty
\newline \\
  \\
  \hline
  (\seq{a}, w, \algo ~ (\epsilon, \mu, \seq{wc}, k)) \leadsto (\seq{a}, w, k) \\
%
% Al-instr
\newline \\
  \seq{a}, w, (\seq{i_1}, \mu, \seq{wc}, k) \vdash i_0 \Rightarrow (w', k') \\
  \hline
  (\seq{a}, w, \algo ~ (i_0 ~ \seq{i_1}, \mu, \seq{wc}, k)) \leadsto (\seq{a}, w', k') \\
%
% Return
\newline \\
  \mu = \getenv(c) \qquad c' = \setenv(c, \mu[s_{var} \mapsto v]) \\
  \hline
  (\seq{a}, w, \ret ~ (v, \call ~ (s_{var}, s_{name}, \seq v, \algo ~ c)) \leadsto
  (\seq a, w, \algo ~ c') \\
\end{gather*}

The semantics of AL is defined by a state transition system.
% TopLevelCall
Given algorithms $\seq a$ generated from the DSL and a Wasm state $w$, a
top-level function $s$ can be invoked with arguments $\seq v$ to perform
either module instantiation or function invocation:
$(\seq a, w, \toplevelcall ~ (s, \seq v))$.
$\toplevelcall ~ (s, \seq v)${} transitions to \algo{} with the function name
$s$, the arguments $\seq v$, and \mt{}.
% Empty
\mt{} is an empty continuation, representing the end of the transition, so no
further transition occurs.
% Call
Similarly to the top-level call, $\call~ (s_{var}, s_{name}, \seq v, k)${}
transitionsto \algo{} with the function name $s_{name}$ and the arguments $\seq
v$, but with the current continuation nested within it.
% Execute
$\exe ~ (\seq{wi}, k)${} executes each Wasm instruction in the sequence
$\seq{wi}$ one by one, each containing a relation name and arguments.
The continuation transitions to \algo{} with the relation name, the arguments,
and the current continuation, with the Wasm instruction being popped.
% Wasm
In $\wasm ~ (\seq{wc}, k)${}, a Wasm instruction is popped from
the Wasm instruction stack and executed until the Wasm context sequence
$\seq{wc}$ is exhausted.
This execution transitions the continuation to \algo{}, with the relation name
and the arguments of the Wasm instruction, and the current continuation nested
within it.
% Algo
In $\algo ~ (\seq i, \mu, \seq{wc}, k)${}, the body instructions $\seq i$
execute sequentially, allowing transitions to \call{}, \exe{}, \wasm{}, \ret{},
or \algo{}.
% Return
The execution of a function call concludes with a return instruction, which
changes the continuation to $\ret ~ v${}.
If the function call originates from \call{}, the return value $v$ is assigned
to the variable $s_{var}$.
If it originates from \toplevelcall{}, the inner continuation $k$ is always
\mt, so the entire execution concludes with the return value $v$:
$
(\seq a, w, \toplevelcall ~ (s, \seq v))
\leadsto^*
(\seq a, w', \ret ~ (v, \mt))
$.

%% Instruction

\begin{gather*}
  \boxed{\seq{a}, ~ w, ~ c \vdash i \Rightarrow w, ~ k} \\
%
% If-true
\newline \\
  w, \getenv(c) \vdash e \Rightarrow v \qquad
  \istrue(v) \\
  \hline
  \seq{a}, w, c \vdash \ifi ~ (e, \seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_1}))) \\
%
% If-false
\newline \\
  w, \getenv(k) \vdash e \Rightarrow v \qquad
  \neg \istrue(v) \\
  \hline
  \seq{a}, w, c \vdash \ifi ~ (e, \seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_2}))) \\
%
% Either-1
\newline \\
  \\
  \hline
  \seq{a}, w, c \vdash \eitheri ~ (\seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_1}))) \\
%
% Either-2
\newline \\
  \\
  \hline
  \seq{a}, w, c \vdash \eitheri ~ (\seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_2}))) \\
%
% Enter
\newline \\
  (\seq{we}, \seq{wi}, sto) = w \qquad
  (i^*, \mu, wc_1 ~ ... ~ wc_n, k) = c \\
  w, \mu \vdash e_1 \Rightarrow wc \qquad
  w, \mu \vdash e_2 \Rightarrow \listv ~ \seq{wi_1} \\
  wi = \getendinstr(wc) \qquad
  wi_1 = \getendinstr(wc_1) \quad ... \quad wi_n = \getendinstr(wc_n) \\
  wi_2^* = wi_1^* ~ wi ~ wi_1 ~ ... ~ wi_n \\
  \hline
  \seq{a}, w, c \vdash \enteri ~ (e_1, e_2)
  \Rightarrow
  (
    (wc ~ \seq{we}, wi_2^* ~ \seq{wi}, sto),
    \wasm ~ (wc ~ wc_1 ~ ... ~ wc_n, \algo ~ (i^*, \mu, \epsilon, k))
  ) \\
%
% PushCtx
\newline \\
  w, \getenv(c) \vdash e \Rightarrow wc \\
  \hline
  \seq{a}, w, c \vdash \pushctxi ~ e
  \Rightarrow
  (\push(w, wc), \algo ~ (\addctx(c, wc))) \\
%
% Push
\newline \\
  w, \getenv(c) \vdash e \Rightarrow wv \\
  \hline
  \seq{a}, w, c \vdash \pushi ~ e \Rightarrow (\push(w, wv), \algo ~ c) \\
%
% PopCtx
\newline \\
  (\seq{we}, \seq{wi}, sto) = w \qquad
  (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \\
  \mu = \getenv(c) \cdot \assign(e, wc) \qquad
  \seq{wi'} = \exit(\seq{wi}) \qquad
  c' = \popwasmctx_{C}(c) \\
  \hline
  \seq{a}, w, c \vdash \popctxi ~ e
  \Rightarrow
  ((\seq{we'}, \seq{wi'}, sto), \algo ~ (\setenv(c', \mu))) \\
%
% Pop
\newline \\
  (w', wv) = \pop(w) \qquad
  \mu' = \getenv(c) \cdot \assign(e, wv) \\
  \hline
  \seq{a}, w, c \vdash \popi ~ e \Rightarrow (w', \algo ~ (\setenv(c, \mu'))) \\
%
% PopN
\newline \\
  \mu = \getenv(c) \qquad
  w, \mu \vdash e_2 \Rightarrow \numv ~ n \\
  (w', wv^n) = \popn(w, n) \qquad
  \mu' = \mu \cdot \assign(e_1, \listv ~ wv^n) \\
  \hline
  \seq{a}, w, c \vdash \popni ~ (e_1, e_2) \Rightarrow
  (w', \algo ~ (\setenv(c, \mu'))) \\
%
% PopAll
\newline \\
  (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \qquad
  \mu' = \getenv(c) \cdot \assign(e, \liste ~ \seq{wv}) \\
  \hline
  \seq{a}, (\seq{we}, \seq{wi}, sto), c \vdash \popalli ~ e
  \Rightarrow
  ((wc ~ \seq{we'}, \seq{wi}, sto), \algo ~ (\setenv(c, \mu'))) \\
%
% Let
\newline \\
  \mu = \getenv(c) \qquad
  w, \mu \vdash e_2 \Rightarrow v \qquad
  \mu' = \mu \cdot \assign(e_1, v) \\
  \hline
  \seq{a}, w, c \vdash \leti ~ (e_1, e_2)
  \Rightarrow
  (w, \setenv(c, \mu')) \\
%
% Trap
\newline \\
  \\
  \hline
  \seq{a}, w, c \vdash \trapi \Rightarrow (w, \ret ~ (\trapv, \mt)) \\
%
% Return-rule
\newline \\
  \\
  \hline
  \seq{a}, w, (\seq{i}, \mu, wc^*, k) \vdash \returnreli \Rightarrow (w, k) \\
%
% Return-func
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  \seq{a}, w, (\seq{i}, \mu, wc^*, k) \vdash \returnfuni ~ e \Rightarrow
  (w, \ret ~ (v, k)) \\
%
% Execute
\newline \\
  w, \getenv(c) \vdash e \Rightarrow \listv ~ wi^* \\
  \hline
  \seq{a}, w, c \vdash \executei ~ e \Rightarrow
  (w, \exe ~ (wi^*, \algo ~ c)) \\
%
% Call-fname
\newline \\
  \mu = \getenv(c) \qquad
  \fnamev ~ s_3 = \mu(s_2) \qquad
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  \seq{a}, w, c \vdash \calli ~ (s_1, s_2, e_1 ~ ... ~ e_n) \Rightarrow
  (w, \call ~ (s_1, s_3, v_1 ~ ... ~ v_n, \algo ~ c)) \\
%
% Call-fname
\newline \\
  \mu = \getenv(c) \qquad
  s_2 \not\in \domain(\mu) \qquad
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  \seq{a}, w, c \vdash \calli ~ (s_1, s_2, \seq e) \Rightarrow
  (w, \call ~ (s_1, s_2, v_1 ~ ... ~ v_n, \algo ~ c)) \\
%
% Replace-frame
\newline \\
\text{NOTE: $e_1$ is either store or current frame} \\
  \mu = \getenv(c) \qquad
  \casev ~ (\text{``Frame"}, v_1 ~ v_2) = \getcurframe(w) \qquad
  w, \mu \vdash e \Rightarrow v \\
  wc = \casev ~ (\text{``Frame"}, v_1 ~ \update(w, \mu, v_2, p^+, v)) \\
  \hline
  \seq{a}, w, c \vdash \replaceframei ~ (p^+, e) \Rightarrow
  (\setcurframe(w, wc), \algo ~ c) \\
%
% Replace-store
\newline \\
  \mu = \getenv(c) \qquad
  w, \mu \vdash e \Rightarrow v \\
  \strv ~ sto = \update(w, \mu, \strv ~ (\getstore(w)), p^+, v) \\
  \hline
  \seq{a}, w, c \vdash \replacestorei ~ (p^+, e)
  \Rightarrow (\setstore(w, sto), c) \\
\end{gather*}

Given algorithms $\seq a$, a Wasm state $w$, and a context $c$, executing an
instruction $i$ produces an updated Wasm state and a continuation.
% If
$\ifi ~ (e, i_1^*, i_2^*)${} evaluates $e$ and branches to $\seq i_1$ if the
result is true, or $\seq i_2$ otherwise.
% Either
$\eitheri ~ (i_1^*, i_2^*)${} nondeterministically selects either $i_1^*$ or
$i_2^*$.
% Enter
$\enteri ~ (e_1, e_2)${} evaluates $e_1$, pushing the resulting Wasm context
$wc$ to the interleaved stack.
It then evaluates $e_2$ to obtain the Wasm instruction sequence $wi_1^*$.
Then, the end instructions of the Wasm context $wc$ and the Wasm contexts
$wc_1 ~ ... ~ wc_n$ in the context $c$ are appended to $wi_1^*$ and pushed to
the Wasm instruction stack.
Finally, it changes the continuation to
$\wasm ~ (wc ~ wc_1 ~ ... ~ wc_n, \algo ~ (i^*, \mu, \epsilon, k))${}.
% PushCtx
$\pushctxi ~ e${} evaluates $e$, pushing the resulting Wasm context $wc$ to the
interleaved stack and the context $c$.
% Push
$\pushi ~ e${} evaluates $e$, pushing the resulting value $wv$ to the
interleaved stack.
% PopCtx
$\popctxi ~ e${} pops all Wasm values up to and including a Wasm context $wc$
from the interleaved stack and all Wasm instructions up to and including an end
instruction from the Wasm instruction stack.
It also pops a Wasm context from the context $c$.
Then, it assign the Wasm context $wc$ to $e$.
% Pop
$\popi ~ e${} pops a Wasm value $wv$ from the interleaved stack and assign the
value to $e$.
% PopN
$\popni ~ (e_1, e_2)${} evaluates $e_2$ to obtain a number $n$, pops $n$ Wasm
values from the interleaved stack, and assigns the values to $e_1$.
% PopAll
$\popalli ~ e${} pops all the Wasm values up to, but not including, a
Wasm context from the interleaved stack and assign the values to $e$.
% Let
$\leti ~ (e_1, e_2)${} evaluates $e_2$ and assigns the resulting value to
$e_1$.
% Trap
\trapi{} results in $\ret ~ (\trapv{}, \mt{})$ to terminate the execution.
% ReturnRel
\returnreli{} results in the continuation $k$ in the context $c$, indicating a
return from a relation algorithm.
% ReturnFunc
$\returnfuni ~ e${} evaluates $e$ to $v$, resulting in $\ret ~ (v, k)${} where
$k$ is the continuation in the context $c$.
% Execute
\executei{} evaluates $e$ to obtain a Wasm instruction sequence $wi^*$ and
result in $\exe ~ (wi^*, \algo ~ c)${}.
% Call
$\calli ~ (s_1, s_2, e_1 ~ ... ~ e_n)${} checks if $s_2$ is bound in the
environment.
If it is, it retrieves the function name $s_3$ from the environment $\mu$;
otherwise, it uses $s_2$ as the function name.
It evaluates $\seq e$ to $v^*$, resulting in $\call ~ (s_1, s_{name}, v_1 ~ ...
~ v_n)${} where $s_{name}$ is the function name.
% ReplaceFrame
\replaceframei{} retrieves the current frame $\casev ~ (\text{``Frame"}, v_1
v_2)${} in the Wasm context $w$.
It then evaluates $e$ to $v$, updating the value at location $v_2$ along paths
$p^+$ to $v$.
% ReplaceFrame
\replacestorei{} evaluates $e$ to $v$, updating the value at location $\strv ~
sto$ along paths $p^+$ to v.

%% Expression

\begin{gather*}
  \boxed{w, \mu \vdash e \Rightarrow v} \\
%
% Var
\newline \\
  v = \mu(s) \\
  \hline
  w, \mu \vdash \vare ~ s \Rightarrow v \\
%
% Num
\newline \\
  \hline
  w, \mu \vdash \nume ~ n \Rightarrow \numv ~ n \\
%
% Bool
\newline \\
  \hline
  w, \mu \vdash \boole ~ b \Rightarrow \boolv ~ b \\
%
% Fname
\newline \\
  \hline
  w, \mu \vdash \fnamee ~ s \Rightarrow \fnamev ~ s \\
%
% List
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \liste ~ e_1 ~ ... ~ e_n \Rightarrow \listv ~ v_1 ~ ... ~ v_n \\
%
% Str
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \stre ~ (s_1, e_1) ~ ... ~ (s_n, e_n) \Rightarrow
  \strv ~ (s_1, v_1) ~ ... ~ (s_n, v_n) \\
%
% Tup
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \tupe ~ e_1 ~ ... ~ e_n \Rightarrow \tupv ~ v_1 ~ ... ~ v_n \\
%
% Case
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \casee ~ (s, e_1 ~ ... ~ e_n) \Rightarrow \casev ~ (s, v_1 ~ ... ~ v_n) \\
%
% Un
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  w, \mu \vdash \une ~ (unop, e) \Rightarrow \unop(unop, v) \\
%
% Bin
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad w, \mu \vdash e_2 \Rightarrow v_2 \\
  \hline
  w, \mu \vdash \bine ~ (binop, e_1, e_2) \Rightarrow \binop(binop, v_1, v_2) \\
%
% Acc
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  w, \mu \vdash \acce ~ (e, p) \Rightarrow \access(w, \mu, v, p) \\
%
% Upd
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad w, \mu \vdash e_2 \Rightarrow v_2 \\
  \hline
  w, \mu \vdash \upde ~ (e_1, p^+, e_2) \Rightarrow \update(w, \mu, v_1, p^+, v_2) \\
%
% Cat
\newline \\
   w, \mu \vdash e_1 \Rightarrow \listv ~ \seq{v_1} \qquad
   w, \mu \vdash e_2 \Rightarrow \listv ~ \seq{v_2} \\
  \hline
  w, \mu \vdash \cate ~ (e_1, e_2) \Rightarrow \listv ~ (\seq{v_1} ~ \seq{v_1}) \\
%
% Comp
\newline \\
   w, \mu \vdash e_1 \Rightarrow \strv ~ \seq{(s_1, v_1)} \qquad
   w, \mu \vdash e_2 \Rightarrow \strv ~ \seq{(s_2, v_2)} \\
  \hline
  w, \mu \vdash \compe ~ (e_1, e_2) \Rightarrow \strv ~ (\seq{(s_1, v_1)} ~ \seq{(s_2, v_2)}) \\
%
% Mem
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad
  w, \mu \vdash e_2 \Rightarrow \listv ~ \seq{v_2} \\
  \hline
  w, \mu \vdash \meme ~ (e_1, e_2) \Rightarrow \boolv ~ (v_1 \in \seq{v_2}) \\
%
% Choose
\newline \\
  w, \mu \vdash e \Rightarrow \listv ~ \seq{v} \qquad
  v \in \seq{v} \\
  \hline
  w, \mu \vdash \choosee ~ e \Rightarrow v \\
%
% Len
\newline \\
  w, \mu \vdash e \Rightarrow \listv ~ \seq{v} \\
  \hline
  w, \mu \vdash \lene ~ e \Rightarrow \numv ~ |\seq v| \\
%
% IsCaseOf
\newline \\
  w, \mu \vdash e \Rightarrow \casev ~ (s', \seq{v}) \\
  \hline
  w, \mu \vdash \iscaseofe ~ (e, s) \Rightarrow \boolv ~ (s = s') \\
%
% GetCurCtx
\newline \\
  wc = \getcurctx(w) \\
  \hline
  w, \mu \vdash \getcurctxe \Rightarrow wc \\
%
% CtxKind
\newline \\
  \casev (s', \seq v) = \getcurctx(w) \\
  \hline
  w, \mu \vdash \ctxkinde ~ s \Rightarrow \boolv ~ (s = s') \\
%
% Iter-
\newline \\
  \,~\mu(s_1) = \listv ~ (v_{11} ~ ... ~ v_{1n}) \quad ... \quad
  \mu(s_m) = \listv ~ (v_{m1} ~ ... ~ v_{mn}) \\
  \mu_1 = \mu[s_1 \mapsto v_{11}, ~ ... ~ s_m \mapsto v_{m1}] \quad ... \quad
  \mu_n = \mu[s_1 \mapsto v_{1n}, ~ ... ~ s_m \mapsto v_{mn}] \\
  w, \mu_1 \vdash e \Rightarrow v_1 \quad ... \quad
  w, \mu_n \vdash e \Rightarrow v_n \\
  \hline
  w, \mu \vdash \itere ~ (e, \listiter, s_1 ~ ... ~ s_m) \Rightarrow
  \listv ~ v_1 ~ ... ~ v_n \\
%
% Iter-n
\newline \\
  w, \mu \vdash e_2 \Rightarrow \numv ~ n \\
  \,~\mu(s_1) = \listv ~ (v_{11} ~ ... ~ v_{1n}) \quad ... \quad
  \mu(s_m) = \listv ~ (v_{m1} ~ ... ~ v_{mn}) \\
  \mu_1 = \mu[s_1 \mapsto v_{11}, ~ ... ~ s_m \mapsto v_{m1}] \quad ... \quad
  \mu_n = \mu[s_1 \mapsto v_{1n}, ~ ... ~ s_m \mapsto v_{mn}] \\
  w, \mu_1 \vdash e \Rightarrow v_1 \quad ... \quad
  w, \mu_n \vdash e \Rightarrow v_n \\
  \hline
  w, \mu \vdash \itere ~ (e_1, \listniter ~ e_2, s_1 ~ ... s_m) \Rightarrow
  \listv ~ v_1 ~ ... ~ v_n \\
%
% Iter-idx
\newline \\
  w, \mu \vdash e_2 \Rightarrow \numv ~ n \\
  \mu_1 = \mu[s_1 \mapsto \numv ~ 0, ~ ... ~ s_m \mapsto \numv ~ 0] \quad ... \quad
  \mu_n = \mu[s_1 \mapsto \numv ~ n, ~ ... ~ s_m \mapsto \numv ~ n] \\
  w, \mu_1 \vdash e \Rightarrow v_1 \quad ... \quad
  w, \mu_n \vdash e \Rightarrow v_n \\
  \hline
  w, \mu \vdash \itere ~ (e_1, \listidxiter ~ e_2, s_1 ~ ... ~ s_m) \Rightarrow
  \listv ~ v_1 ~ ... ~ v_n \\
%
% Match
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad
  w, \mu \vdash e_2 \Rightarrow v_2 \\
  \hline
  w, \mu \vdash \matche ~ (e_1, e_2) \Rightarrow \boolv ~ (\match(v_1, v_2)) \\
%
% HasType
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  w, \mu \vdash \hastypee ~ (e, s) \Rightarrow \boolv ~ (\hastype(v, s)) \\
\end{gather*}

Given a Wasm state $w$ and an environment $\mu$, an expression $e$ evaluates
to a value.
% Var
$\vare ~ s${} looks up the environment $\mu$ and gets the value.
% Num Bool Fname
$\nume ~ n${}, $\boole ~ b${}, and $\fnamee ~ s${} directly reduce to $\numv ~
n${}, $\boolv ~ b${}, and $\fnamev ~ s${}, respectively.
% Tuple Case List Struct
$\liste ~ e_1 ~ ... ~ e_n${}, $\stre ~ (s_1, e_1) ~ ... ~ (s_n, e_n)${}, $\tupe
~ e_1 ~ ... ~ e_n${}, and $\casee ~ (s, e_1 ~ ... ~ e_n)${} evaluate their
sub-expressions and reduce to $\listv ~ v_1 ~ ... ~ v_n${}, $\strv ~ (s_1, v_1)
~ ... ~ (s_n, v_n)${}, $\tupv ~ v_1 ~ ... ~ v_n${}, and $\casev ~ (s, v_1 ~ ...
~ v_n)${}, respectively.
% Un Bin
$\une ~ (unop, e)${} and $\bine ~ (binop, e_1, e_2)${} evaluate their
sub-expressions and perform \unop{} and \binop{}, respectively.
% Acc
$\acce ~ (e, p)${} evaluates $e$ and accesses the resulting value $v$ using a
path $p$.
% Upd
$\upde ~ (e_1, p^+, e_2)${} evaluates $e_1$ and $e_2$ to $v_1$ and $v_2$,
updating the value at location $v_1$ along paths $p^+$ to the value $v_2$.
% Cat Comp
$\cate ~ (e_1, e_2)${} and $\compe ~ (e_1, e_2)${} each evaluate two
sub-expressions $e_1$ and $e_2$, concatenating the resulting lists and
resulting records, respectively.
% Mem
$\meme ~ (e_1, e_2)${} evaluates $e_1$ and $e_2$ to $v_1$ and $\listv ~ \seq
v_2$, and checks whether $v_1$ is a member of $\seq v_2$.
% Choose Len
$\choosee ~ e$ and $\lene ~ e${} each evaluate $e$ to the list value $\listv ~
\seq v$, choosing an element from the list and calculating the length of the
list, respectively.
% IsCaseOf
$\iscaseofe ~ (e, s)${} evaluates $e$ to the tagged tuple $\casev ~ (s', \seq
v)$ and checks whether the tag $s'$ matches $s$.
% GetCurCtx
\getcurctxe{} retrieves the current Wasm context from the interleaved stack.
% CtxKind
$\ctxkinde ~ s${} retrieves the current Wasm context $\casev ~ (s', \seq v)$
and checks whether the kind $s'$ matches $s$.
% Iter
$\itere ~ (e, iter, s^*)${} behaves differently depending on $iter$.
\listiter{} and $\listniter ~ e_2${} are essentially equivalent.
They assume that all the values mapped to the variables $\seq s$ by the
environment $\mu${} are list values with the same length $n$.
$\listniter ~ e_2${} checks whether the number $n$ matches the result of
evaluating $e_2$, while \listiter{} allows any number.
It generates a sequence of environments $\mu_1 ~ ... ~ \mu_n$ where the k-th
environment maps each variable to the k-th elements of each list.
Each environment is used to evaluate $e_1$, producing a sequence of values $v_1
~ ... ~ v_n$.
Finally, it results in $\listv ~ v_1 ~ ... ~ v_n$.
$\listidxiter ~ e_2${} behaves similarly except that the variables $\seq s$ are
unbound.
It evaluates $e_2$ to obtain $n$ and then behaves like $\listniter ~ e_2${} as
if a list of values [0 ... n-1] were bound to the variables $s^*$.
% IsValid Match HasType
$\matche ~ (e_1, e_2)${} and $\hastypee ~ (e, s)${} evaluate their
sub-expressions and perform \match{} and \hastype{}, respectively.
Note that \match{} and \hastype{} represent the match relation and the type
relation in the Wasm static semantics.
Their formalization is omitted as they are highly specific to Wasm and
excessively verbose.

%% Helper functions

\begin{align*}
%
% get_algo_name
\newline \\
  &\getalgoname(\rel ~ (s, \seq e, \seq i)) = s \\
  &\getalgoname(\fun ~ (s, \seq e, \seq i)) = s \\
%
% lookup
\newline \\
  &\lookup(\epsilon, s) = \epsilon \\
  &\lookup(a ~ \seq{a}, s) =
    \begin{cases}
      a &\quad\quad \premise{if}\quad ~ s = \getalgoname(a) \\
      \lookup(\seq{a}, s) &\quad\quad \premise{otherwise}
    \end{cases}
  \\
%
% assign
\newline \\
  &\assign(\vare ~ s, v) = [s \mapsto v] \\
  &\assign(\itere ~ (e, iter, s^*), v) = \assign(e, v) \\
  &\assign(\tupe ~ e_1 ~ ... ~ e_n, \tupv ~ v_1 ~ ... ~ v_n) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(\liste ~ e_1 ~ ... ~ e_n, \listv ~ v_1 ~ ... ~ v_n) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(\casee ~ (s, e_1 ~ ... ~ e_n), \casev ~ (s, v_1 ~ ... ~ v_n)) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(
    w,
    \mu,
    \stre ~ (s_1, e_1) ~ ... ~ (s_2, e_n),
    \strv ~ (s_1, v_1) ~ ... ~ (s_n, v_n)
  ) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(\cate ~ (\liste ~ e_1 ~ ... ~ e_m, e), \listv v_1 ~ ... ~ v_n) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_m, v_m) \cdot
    \assign(e, \listv ~ v_{m+1} ~ ... ~ v_n) \\
  &\assign(\cate ~ (e, \liste ~ e_1 ~ ... ~ e_m), \listv v_1 ~ ... ~ v_n) =
    \assign(e, \listv ~ v_1 ~ ... ~ v_{n-m}) \cdot
    \assign(e_1, v_{n-m+1}) \cdot ~ ... ~ \cdot \assign(e_m, v_n) \\
%
% create_algo
\newline \\
  &\createalgo(\seq a, s, \seq v, k) =
  \algo(\seq i, \mu, \epsilon, k) \\
  &\qquad\qquad\qquad\qquad\premise{if}\quad
  (\rel ~ (s, \seq e, \seq i) \quad\lor\quad \fun ~ (s, \seq e, \seq i)) = \lookup(\seq a, s)
  \quad\land\quad
  \mu = \assign(w, [\text{``s"} \mapsto \storev], \seq e, \seq v) \\
%
% split_ctx
\newline \\
  &\splitctx(\epsilon) = (\epsilon, \epsilon, \epsilon) \\
  &\splitctx(we ~ \seq{we}) =
    \begin{cases}
      (\epsilon, wc, \seq{we})
      &\quad\quad \premise{if}\quad
      we = wc \\
      (wv ~ \seq{wv}, wc, \seq{we'})
      &\quad\quad \premise{if}\quad
      we = wv \quad\land\quad
      (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \\
    \end{cases}
  \\
%
% pop-wasm-ctx_context
\newline \\
  &\popwasmctx_{C}((\seq i, \mu, epsilon, k)) = (\seq i, \mu, \epsilon, \popwasmctx_K(k)) \\
  &\popwasmctx_{C}((\seq i, \mu, wc ~ \seq{wc}, k)) = (\seq i, \mu, wc^*, k) \\
%
\newline \\
% pop-wasm-ctx_cont-mt
  &\popwasmctx_{K}(\mt) = \epsilon \\
% pop-wasm-ctx_cont-toplevelcall
  &\popwasmctx_{K}(\toplevelcall ~ (s, v^*)) = \epsilon \\
% pop-wasm-ctx_cont-call
  &\popwasmctx_{K}(\call ~ (s_1, s_2, \seq v, k)) =
    \call ~ (s_1, s_2, \seq v, (\popwasmctx_K(k)) \\
% pop-wasm-ctx_cont-execute
  &\popwasmctx_{K}(\exe ~ (v, k)) = \exe ~ (v, (\popwasmctx_K(k)) \\
% pop-wasm-ctx_cont-wasm_0
  &\popwasmctx_{K}(\wasm ~ (\epsilon, k)) = \wasm ~ (\epsilon, \popwasmctx_{K}(k)) \\
% pop-wasm-ctx_cont-wasm_nonzero
  &\popwasmctx_{K}(\wasm ~ (wc ~ \seq{wc}, k)) = \wasm ~ (\seq{wc}, k) \\
% pop-wasm-ctx_cont-al
  &\popwasmctx_{K}(\algo ~ c) = \algo ~ (\popwasmctx_C(c)) \\
% pop-wasm-ctx_cont-ret
  &\popwasmctx_{K}(\ret ~ (v, k)) = \ret ~ (v, \popwasmctx_K(k)) \\
%
% exit
\newline \\
  &\exit(wi ~ \seq{wi}) =
    \begin{cases}
      \seq{wi} &\quad\quad \premise{if}\quad ~ \isendinstr(wi) \\
      \exit(\seq{wi}) &\quad\quad \premise{otherwise} \\
    \end{cases}
  \\
%
% get_env
\newline \\
  &\getenv((\seq i, \mu, wc^*, k)) = \mu \\
%
% set_env
\newline \\
  &\setenv((\seq i, \mu, wc^*, k), \mu') = (\seq i, \mu', wc^*, k) \\
%
% prepend_instr
\newline \\
  &\prependinstr((\seq i, \mu, wc^*, k), \seq{i'}) = (\seq{i'} ~ \seq i, \mu, wc^*, k) \\
%
% add_ctx
\newline \\
  &\addctx((\seq i, \mu, wc^*, k), wc') = (\seq i, \mu, wc' ~ wc^*, k) \\
%
% get_store
\newline \\
  &\getstore((\seq{we}, \seq{wi}, sto)) = sto \\
%
% set_store
\newline \\
  &\setstore((\seq{we}, \seq{wi}, sto), sto') = (\seq{we}, \seq{wi}, sto') \\
%
% pop_wasm_instr
\newline \\
  &\popwasminstr((\seq{we}, wi ~ \seq{wi}, sto)) = ((\seq{we}, \seq{wi}, sto), wi) \\
%
% push
\newline \\
  &\push((\seq{we}, \seq{wi}, sto), we) = (we ~ \seq{we}, \seq{wi}, sto) \\
%
% pop
\newline \\
  &\pop((we ~ \seq{we}, \seq{wi}, sto)) = ((\seq{we}, \seq{wi}, sto), we) \\
%
% popn
\newline \\
  &\popn((we^n ~ \seq{we}, \seq{wi}, sto), n) = ((\seq{we}, \seq{wi}, sto), we^n) \\
%
% unop
\newline \\
  &\unop(\notop, v) =
    \begin{cases}
      \boolv ~ true &\quad\quad\premise{if}\quad ~ \istrue(v) \\
      \boolv ~ false &\quad\quad\premise{otherwise} \\
    \end{cases}
  \\
  &\unop(\minusop, \numv ~ n) = \numv ~ (-n) \\
%
% get_cur_ctx
\newline \\
  &\getcurctx((\seq{we}, \seq{wi}, sto)) =
  wc \quad\quad\premise{if}\quad ~ (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \\
%
% get_cur_frame-frame
\newline \\
  &\getcurframe((\seq{we}, \seq{wi}, sto)) =
    \begin{cases}
      wc &\quad\quad\premise{if}\quad
      (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad\land\quad \isframe(wc)\\
      \getcurframe(\seq{we'}) &\quad\quad\premise{if}\quad
      (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad\land\quad \neg \isframe(wc) \\
    \end{cases}
  \\
%
% set_cur_frame-frame
\newline \\
  &\setcurframe((\seq{we}, \seq{wi}, sto), wc_{frame})
  =
  \begin{cases}
    (\seq{wv} ~ wc_{frame} ~ \seq{we'}, \seq{wi}, sto) \\
    \qquad\qquad\qquad\quad\premise{if}\quad
    (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad \isframe(wc) \\
    (\seq{wv} ~ wc ~ \setcurframe(\seq{we'}, wc_{frame}), \seq{wi}, sto) \\
    \qquad\qquad\qquad\quad\premise{if}\quad
    (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad \neg \isframe(wc) \\
  \end{cases}
  \\
%
% access-idx
\newline \\
  &\access(w, \mu, \listv ~ \seq{v}, \idxp ~ e) = \seq{v}[n]
  \qquad\qquad\qquad~\,\quad\premise{if}\quad w, \mu \vdash e \Rightarrow \numv ~ n \\
% access-slice
  &\access(w, \mu, \listv ~ \seq{v}, \slicep ~ (e_1, e_2)) = \seq{v}[n_1: n_2]
  \quad\quad\premise{if}\quad
  w, \mu \vdash e_1 \Rightarrow \numv ~ n_1 \quad\land\quad
  w, \mu \vdash e_2 \Rightarrow \numv ~ n_2 \\
% access-dot
  &\access(w, \mu, \strv ~ ((s_0, v_0) ~ \seq{(s_1, v_1)}), \dotp ~ s) =
  \begin{cases}
    v_0 &\quad\quad\premise{if}\quad s_0 = s \\
    \access(w, \mu, \strv ~ \seq{(s_1, v_1)}, \dotp ~ s) &\quad\quad\premise{otherwise} \\
  \end{cases}
  \\
% access-store
  &\access(w, \mu, \storev, p) =\access(w, \mu, \strv ~ sto, p)
  \quad\quad\premise{if}\quad w = (\seq{we}, \seq{wi}, sto)\\
%
% update
\newline \\
  % idx
  &\update(w, \mu, \listv ~ \seq{v}, (\idxp ~ e) ~ \seq{p}, v) =
  \listv ~ (\updateidx(\seq v, n, \update(w, \mu, \seq{v}[n], \seq{p}, v))) \\
  &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\premise{if}\quad
  w, \mu \vdash e \Rightarrow \numv ~ n \\
  % slice
  &\update(w, \mu, \listv ~ \seq{v}, (\slicep ~ e) ~ \seq{p}, v) =
  \listv ~ (\updateslice(\seq v, n_1, n_2, \update(w, \mu, \seq{v_i}[n_1: n_2], \seq{p}, v))) \\
  &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\premise{if}\quad
  w, \mu \vdash e_1 \Rightarrow \numv ~ n_1 \quad\land\quad
  w, \mu \vdash e_2 \Rightarrow \numv ~ n_2 \\
  % dot
  &\update(w, \mu, \strv ~ \seq{(s, v)}, (\dotp ~ s') ~ \seq{p}, v)
  =
  \strv ~ (\updatedot(\seq{(s, v)}, s', \update(w, \mu, \access(v, \dotp ~ s'), \seq{p}, v))) \\
%
% update_idx
\newline \\
  &\updateidx(v_0 ~ \seq{v_1}, n, v_2) =
  \begin{cases}
    v_2 ~ \seq{v_1}
    &\quad\quad\premise{if}\quad n = 0 \\
    v_0 ~ \updateidx(\seq{v_1}, n-1, v_2)
    &\quad\quad\premise{if}\quad n > 0 \\
  \end{cases} \\
%
% update_slice
\newline \\
  &\updateslice(v_0^n ~ \seq{v_1}, 0, n, v_2^n) = v_2^n ~ \seq{v_1} \\
  &\updateslice(v_0 ~ \seq{v_1}, m, n, v_2^n) =  v_0 ~ \updateslice(\seq{v_1}, m-1, n, v_2)
  \quad\quad\premise{if}\quad m > 0 \\
%
% update_dot
\newline \\
  &\updatedot((s_0, v_0) ~ \seq{(s_1, v_1)}, s_2, v_2) =
  \begin{cases}
    (s_2, v_2) ~ \seq{(s_1, v_1)}
    &\quad\quad\premise{if}\quad s_0 = s_2 \\
    (s_0, v_0) ~ \updatedot(\seq{(s_1, v_1)}, s_2, v_2)
    &\quad\quad\premise{otherwise}
  \end{cases}
  \\
%
% get_end_algo
\newline \\
  &\getendinstr(wc) =
  \begin{cases}
    \casev ~ (\text{``Exiting"}, \epsilon) &\quad\quad\premise{if}\quad wc = \casev ~ (\text{``Label"}, \seq v)\\
    \casev ~ (\text{``Returning"}, \epsilon) &\quad\quad\premise{if}\quad wc = \casev ~ (\text{``Frame"}, \seq v)
  \end{cases} \\
%
% is_end_instr
\newline \\
  &\isendinstr(wi) =
  \begin{cases}
    true &\quad\quad\premise{if}\quad wi = \casev ~ (\text{``Exiting"}, \epsilon) \quad\lor\quad wi = \casev ~ (\text{``Returning"}, \epsilon)\\
    false &\quad\quad\premise{otherwise}
  \end{cases} \\
%
% is_true
\newline \\
  &\istrue(v) =
  \begin{cases}
  b &\quad\quad\premise{if}\quad v = \boolv ~ b \\
    \istrue(v_1) ~ \land ~ ... ~ \land ~ \istrue(v_n) &\quad\quad\premise{if}\quad
    v = \listv ~ v_1 ~ ... ~ v_n \\
  \end{cases} \\
%
% is_frame
\newline \\
  &\isframe(v) =
  \begin{cases}
    true
    &\quad\quad\premise{if}\quad v = \casev ~ (\text{"FRAME"}, \seq v) \\
    false &\quad\quad\premise{otherwise}\\
  \end{cases}
  \\
\end{align*}