### 2024-11-22 paper/model.tex
Fix chapter title

\chapter{Algorithmic Specification Language}
\label{ch:Model}
\noindent

\section{Control flow in SpecTec prose}
\label{sec:spectecp}

% TODO: Remove these commands
\newcommand{\officialp}{official prose}
\newcommand{\spectecp}{SpecTec prose}

% control structure
It might be easy to understand how the control flow of Wasm works, if we assume
that a Wasm code is loaded on a memory space and a pc points to the instruction
to execute.
To describe control flow, it introduces a structure named \textit{block} which
consitutes of a instruction sequence.
When branching instruction in the block is performed, pc can be changed to the
starting point of the block or the end of the block.

% an example of Wasm control flow in official prose
\begin{example}
\label{ex:br}
\begin{verbatim}
  // infinite loop
  (loop (result i32) (i32.const 42) (br 0) (unreachable) end) (unreachable)
\end{verbatim}
\end{example}

\cref{ex:br} is a Wasm code example that has \texttt{loop} and
\texttt{unreachable}.
Here, result type of the \texttt{loop} is \texttt{i32} and the \texttt{loop}
has a block of three instructions: \texttt{i32.const}, \texttt{br}, and
\texttt{unreachable}.
\cref{fig:loop} is the specification for the \texttt{loop} instruction in the
prose notation.
In forth step, it says that the continuation is the start of the loop, the
information is stored in a label.
In the last step, it \textit{enters} the block with the label.
The term \textit{enter} means that it pushes the label onto the stack and makes
pc points to the first instruction in the block.
The \texttt{i32.const} instruction is executed first, which just pushes the
\texttt{i32} value \texttt{42} in the stack.
Then, The \texttt{br} instruction is executed.
\cref{fig:br} is the \officialp{} of the \texttt{br} instruction.
It pops the label from the stack, and makes pc points to the continuation of
the label, which is the start of the loop instruction.
As a result, the code example above is a infinite loop pushing \texttt{42}
forever without executing \texttt{unreachable} instructions.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/loop}}
    \caption[Enter the caption title here]{\texttt{loop} instruction} \label{fig:loop}
    \centerline{\includegraphics[width=15cm]{fig/br}}
    \caption[Enter the caption title here]{\texttt{br} instruction} \label{fig:br}
\end{figure}

% exiting label in official prose
There is also a special form of a behavior related to the block: an
\textit{exiting label}.
\cref{fig:exiting-label} is the \officialp{} of the \textit{exiting label}.
It is special because this behavior is performed without an explicit
Wasm instruction.
Rather, the behavior is performed when \textbf{the end of a block is reached}
without control instructions or runtime error.
The behavior is that the label is popped from the stack, and the pc changes to
the point after the end of the block.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/exiting}}
    \caption[Enter the caption title here]{exiting label} \label{fig:exiting-label}
\end{figure}

% an example of exiting label in official prose
\begin{example}
\label{ex:exit}
\begin{verbatim}
  // exiting label
  (loop (result i32) (i32.const 42) end) (f64.const 3.14)
\end{verbatim}
\end{example}

\cref{ex:exit} is a Wasm code example that has \texttt{loop} whose block is
only \texttt{i32.const}, and \texttt{f32.const}
When the \texttt{loop} instruction is executed, it enters the block with a label.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
As a result, after executing the code, the value 42 and 3.14 pushed to the
stack.

% control flow structure with function call in official prose
Similar to the control flow using the block and the label, there is a function
call with a frame.
\cref{fig:invoke} is the \officialp{} of the function invocation which takes
place by a \texttt{call} instruction.
A frame is pushed to the stack and enters the block of the function body with a
label.
There is also a special behavior named returning from a function.
\cref{fig:returning} is the \officialp{} of the returning from a function.
When end of function is reached without without control instructions or runtime
error, The frame is popped from the stack, and the pc points to the next
instruction of the caller instruction.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/invoke}}
    \caption[Enter the caption title here]{function invocation} \label{fig:invoke}
    \centerline{\includegraphics[width=15cm]{fig/returning}}
    \caption[Enter the caption title here]{returning from a function} \label{fig:returning}
\end{figure}

% control flow structure in SpecTec prose
However, the \spectecp{} assumes a different model to explain the control flow
of WebAssembly.
Rather than using the notion of a pc, \spectecp{} assumes that WebAssembly
instructions are just given one by one.
This is because the \spectecp{} is generated automatically from the
\red{SpecTec DSL}.
\red{SpecTec DSL} uses rewrite rule to describe WebAssembly semantics, and it
is interpreted in \spectecp{} as consuming a WebAssembly instruction, pushing
or poping a control structures, and inputing new WebAssembly instructions.
It can be expressed in the following form:
$c^* \vdash i_0, i_1, ..., i_m \leadsto c'^* \vdash i'_0, ...i'_{m'}, i_1, ..., i_m$.
It means that given a sequence of contexts $c^*$, executing $i_0$ results in a
new sequence of contexts $c'^*$ and instructions $i'_0, ...i'_{m'}$.
To simply model the control flow of WebAssembly, other things not closely
related to the control flow are omitted in this notation.

% challenges in SpecTec prose
However, this viewpoint is challenged by the exiting label.
As the block structure doesn't remain intact in this model, it is hard to
describe \textbf{the end of the block}.
Furthermore, the exiting label is not performed by a specific WebAssembly
instruction, which makes it hard to model the behavior.
To handle this problem, an administrative instruction \texttt{end} is
introduced in \spectecp.
When entering a block, an \texttt{end} is appended to the end of the block.

% an example of Wasm control flow in official prose
\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/spectec-loop}}
    \caption[Enter the caption title here]{SpecTec \texttt{loop}} \label{fig:spectec-loop}
\end{figure}
\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/spectec-br}}
    \caption[Enter the caption title here]{SpecTec \texttt{br}} \label{fig:spectec-br}
\end{figure}

\red{TODO: Change to array} \\
\begin{align}
  &c^*
  \vdash
  loop([ ~ i32.const ~ 42, ~ br ~ 0, ~ unreachable ~ ]), ~ unreachable
  \label{eq:loop-1} \\
&\leadsto
  label(loop([ ~ i32.const ~ 42; ~ br ~ 0; ~ unreachable ~ ])), c^*
  \vdash
  i32.const ~ 42, ~ br ~ 0, ~ unreachable, ~ end, ~ unreachable
  \label{eq:loop-2} \\
  &\leadsto
  label(loop([ ~ i32.const ~ 42; ~ br ~ 0; ~ unreachable ~ ])), c^*
  \vdash
  ~ br ~ 0, ~ unreachable, ~ end, ~ unreachable
  \label{eq:loop-3} \\
&\leadsto
  c^*
  \vdash
  loop([ ~ i32.const ~ 42; ~ br ~ 0; ~ unreachable ~ ]), unreachable
  \label{eq:loop-4}
\end{align}

Consider \cref{ex:br} above.
If we assume some control structure $c^*$ is given, the code can be expressed
like \cref{eq:loop-1}.
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself in the label, and \textit{enters} the
block.
Here, \textit{enter} means that it pushes the label in the stack and executes
the instructions in the block, which means that \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, \texttt{end} and \texttt{unreachable} become
the new inputs in \cref{eq:loop-2}.
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does
in \cref{eq:loop-3}.
The value is omitted for brevity.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block including the \texttt{end}, considers the loop instruction in the
label as a new input instruction.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}
again in \cref{eq:loop-4}.
Therefore, it explains the same behavior in a different point of view.

\red{TODO: Also use example 3 to explain function call in official prose}

% an example of function call in official prose
\begin{example}
\label{ex:invoke}
\begin{verbatim}
  // function definition of $push42
  (func $push42 (result i32) (i32.const 42))

  // function call
  (call $push42) (f32.const 3.14)
\end{verbatim}
\end{example}

\red{TODO: Change to array} \\
\begin{align}
  &c^* \vdash call(\$push42), ~ f32.const ~ 3.14 \label{eq:call-1} \\
&\leadsto
  Label(\epsilon), ~ Frame, ~ c^* \vdash i32.const ~ 42, ~ end, ~ f32.const ~ 3.14 \label{eq:call-2} \\
&\leadsto
  Label(\epsilon), ~ Frame, ~ c^* \vdash end, ~ f32.const ~ 3.14 \label{eq:call-3} \\
&\leadsto
  c^* \vdash f32.const ~ 3.14 \label{eq:call-4} \\
&\leadsto
  c^* \vdash \epsilon \label{eq:call-5}
\end{align}

\cref{ex:invoke} contains a function definition whose body is \texttt{i32.const}, a
\texttt{call} instruction followed by a \texttt{f32.const} instruction, which
can be expressed in \cref{eq:call-1}.
By invoking a function, it pushes a frame and enters the block with a label in
\cref{eq:call-2}.
After executing the body in \cref{eq:call-3}, an \texttt{end} is executed.
In this point, this \texttt{end} represent both the end of the block and the
function.
As a consequence, this \text{end} perform both exiting label and returning from
a function in \cref{eq:call-4}, and the instruction after the \texttt{call} is
executed in \cref{eq:call-5}
It results in the values 42 and 3.14 pushed to the stack.

% summary & test
In short, the behavior of an \texttt{end} instruction 1) performs exiting
label, 2) looks up the top of the context, and 3) performs returing from a
function, if it is a frame.
The \spectecp{} can model the WebAssembly control flow using this \texttt{end}
instruction.
To show the correctness of the \spectecp{}, we run the WebAssembly tests
including official WebAssembly test suite by executing the \spectecp{} with
\red{definitional interpreter}.

% Problem
However, tests related to the control instruction are failed.
At that time, there was no explicit model to describe the control flow in
\spectecp{}, it was hard to define the problem, making it complicated to fix
the problem.

%%%%%%%%

% an example of buggy test case
\begin{example}
\label{ex:bug}
\begin{verbatim}
  // function definition of $br-returning
  (func  $brAndReturning (br 0) (unreachable))

  // function call
  (call $brAndReturning)
\end{verbatim}
\end{example}

\cref{ex:bug} contains a function that has only a \texttt{br} instruction, and a
\texttt{call} insstruction.
According to the \officialp{}, when the function \texttt{\$brAndReturning} is
called, a frame is pushed to the stack, and the block of the function
body is entered with a label.
When the \texttt{br} is executed, it pops the label from the stack.
As the end of the function is reached after the execution, returning from a
function is performed.
As a result, the frame is popped, so the call actually does nothing.

% explaination of the bug in SpecTec prose
\begin{align}
  &c^* \vdash call(\$brAndReturning) \label{eq:bug-1} \\
&\leadsto
  Label(\epsilon), ~ Frame, ~ c^* \vdash (br ~ 0) ~ (unreachable), ~ end \label{eq:bug-2} \\
&\leadsto
  Frame, ~ c^* \vdash \epsilon \label{eq:bug-3}
\end{align}

In \spectecp{} the call instruction in \cref{eq:bug-1} pushes a frame and
enters the function body \cref{eq:bug-2}.
When the \texttt{br} is executed, it pops the label and removes the input
instructions until the end in \cref{eq:bug-3}.
As a result, the frame is remained after the executing the call.
The problem is that the \texttt{br} should remove the end of the block but
should not remove the end of the function, but \texttt{end} express the end of
the block and the function at the same time.

% Fix
To fix the problem, semantics of \enteri is changed.
The sequence of contexts that is pushed in the AL algorithm is stored,
and when a block is entered, exiting label and returning from a function is
appended according to the sequence.

\newcommand{\seq}[1]{#1^*}

%% Syntax
\section{Syntax of AL}
\label{syntax}

{
\renewcommand{\arraystretch}{0.8835}  % Decrease row spacing locally
\begin{align*}
\begin{array}{lcccrlr}
%
% Algorithm
  \text{Algorithm}\quad& A &\ni& a &::=& ~ \rel ~ (s, \seq e, \seq i) ~ | ~ \fun ~ (s, \seq e, \seq i) \\
%
% Instruction
  \text{Instruction}\quad& I &\ni& i &::=& ~ \ifi ~ (e, \seq i, \seq i) &\quad\text{(If)} \\
    &&&& | & ~ \eitheri ~ (\seq i, \seq i) &\quad\text{(Either)} \\
    &&&& | & ~ \enteri ~ (e, e) &\quad\text{(Enter)} \\
    &&&& | & ~ \pushctxi ~ e &\quad\text{(Push Context)} \\
    &&&& | & ~ \pushi ~ e &\quad\text{(Push)} \\
    &&&& | & ~ \popctxi ~ e &\quad\text{(Pop Context)} \\
    &&&& | & ~ \popi ~ e &\quad\text{(Pop)} \\
    &&&& | & ~ \popni ~ (e, e) &\quad\text{(Pop N)} \\
    &&&& | & ~ \popalli ~ e &\quad\text{(Pop All)} \\
    &&&& | & ~ \leti ~ (e, e) &\quad\text{(Let)} \\
    &&&& | & ~ \trapi &\quad\text{(Trap)} \\
    &&&& | & ~ \returnreli &\quad\text{(Return Relation)} \\
    &&&& | & ~ \returnfuni ~ e &\quad\text{(Return Function)} \\
    &&&& | & ~ \executei ~ e &\quad\text{(Execute)} \\
    &&&& | & ~ \calli ~ (s, s, \seq{e}) &\quad\text{(Let Call)} \\
    &&&& | & ~ \replaceframei ~ (p^+, e) &\quad\text{(Replace Frame)} \\
    &&&& | & ~ \replacestorei ~ (p^+, e) &\quad\text{(Replace Store)} \\
%
% Expression
  \text{Expression}\quad& E &\ni& e &::=& ~ \vare ~ s &\quad\text{(Variable)} \\
    &&&& | & ~ \nume ~ n &\quad\text{(Number)} \\
    &&&& | & ~ \boole ~ b &\quad\text{(Boolean)} \\
    &&&& | & ~ \fnamee ~ s &\quad\text{(Function Name)} \\
    &&&& | & ~ \liste ~ \seq e &\quad\text{(List)} \\
    &&&& | & ~ \stre ~ \seq{(s, e)} &\quad\text{(Record)} \\
    &&&& | & ~ \tupe ~ \seq e &\quad\text{(Tuple)} \\
    &&&& | & ~ \casee ~ (s, \seq e) &\quad\text{(Tagged Tuple)} \\
    &&&& | & ~ \une ~ (unop, e) &\quad\text{(Unary Operation)} \\
    &&&& | & ~ \bine ~ (binop, e, e) &\quad\text{(Binary Operation)} \\
    &&&& | & ~ \acce ~ (e, p) &\quad\text{(Access)} \\
    &&&& | & ~ \upde ~ (e, p^+, e) &\quad\text{(Update)} \\
    &&&& | & ~ \cate ~ (e, e) &\quad\text{(Concatenation)} \\
    &&&& | & ~ \compe ~ (e, e) &\quad\text{(Composition)} \\
    &&&& | & ~ \meme ~ (e, e) &\quad\text{(Membership)} \\
    &&&& | & ~ \choosee ~ e &\quad\text{(Choose)} \\
    &&&& | & ~ \lene ~ e &\quad\text{(Length)} \\
    &&&& | & ~ \iscaseofe ~ (e, s) &\quad\text{(Check Tag)} \\
    &&&& | & ~ \getcurctxe  &\quad\text{(Get Current Context)} \\
    &&&& | & ~ \ctxkinde ~ s &\quad\text{(Context Kind)} \\
    &&&& | & ~ \itere ~ (e, iter, \seq{s}) &\quad\text{(Iteration)} \\
    &&&& | & ~ \matche ~ (e, e) &\quad\text{(Match)} \\
    &&&& | & ~ \hastypee ~ (e, s) &\quad\text{(Has Type)} \\
\end{array}
\end{align*}
}

\newpage
\begin{align*}
\begin{array}{lcccrlr}
%
% Path
  \text{Path}\quad& P &\ni& p &::=& ~ \idxp ~ e ~ | ~ \slicep ~ (e, e) ~ | ~ \dotp ~ s \\
%
% Iter
  \text{Iter}\quad& Iter &\ni& iter &::=& ~ \listiter ~ | ~ \listniter ~ e ~|~ \listidxiter ~ (s, e) \\
%
% Operator
  \text{Unary Operator}\quad& Unop &\ni& unop &::=& ~ \notop ~ | ~ \minusop \\
  \text{Binary Operator}\quad& Binop &\ni& binop &::=& ~ \addop ~ | ~ \subop ~ | ~ \mulop \\
    &&&& | & \divop ~ | ~ \modop ~ | ~ \expop \\
    &&&& | & \implop ~ | ~ \equivop ~ | ~ \andop \\
    &&&& | & \orop ~ | ~ \eqop ~ | ~ \neop ~ | ~ \ltop \\
    &&&& | & \gtop ~ | ~ \leop ~ | ~ \geop \\
% Primitives
  \text{String}\quad& \mathbb S &\ni& s \\
  \text{Integer}\quad& \mathbb Z &\ni& n \\
  \text{Boolean}\quad& \mathbb B &\ni& b \\
\end{array}
\end{align*}

% algorithm
An algorithm is either an AL relation or an AL function, corresponding to the
relation and the function definitions in the DSL.
% instruction
An AL instruction is \ifi{} for conditionally choosing a branch, \eitheri{} for
nondeterministically choosing a branch, \enteri{} for entering a block,
\pushctxi{} for pushing a Wasm context, \pushi{} for pushing a Wasm value,
\popctxi{} for popping a Wasm context, \popi{} for popping a Wasm value,
\popni{} for popping $n$ Wasm values, \popalli{} for popping all Wasm values
within a Wasm context, \leti{} for let binding, \trapi{} for trapping,
\returnreli{} for returning from an AL relation, \returnfuni{} for returning
from an AL function, \executei{} for executing given Wasm instructions,
\calli{} for an AL function call with let binding, \replaceframei{} for
replacing a frame, or \replacestorei{} for replacing a store.
% expression
An expression is a variable, number, boolean, AL function name, list, record
(or struct), tuple, tagged tuple, unary or binary operation, data structure
access or update, list or record concatenation, membership check, element
selection, length retrieval, tag check, Wasm context retrieval, Wasm context
kind check, iteration, match relation, or type relation.

\begin{align*}
\begin{array}{lcccrlr}
%
% State
  \text{State}\quad& \Sigma &\ni& \sigma & ::=& ~ \seq a, w, k \\
%
% Wasm state
  \text{Wasm State}\quad& W &\ni& w &::=& ~ \seq{we}, \seq{wi}, sto \\
%
% Wasm Entry
  \text{Wasm Entry}\quad& WE &\ni& we &::=& ~ wv ~ | ~ wc \\
%
% Wasm Value
  \text{Wasm Value}\quad& WV &\ni& wv &::=& ~ \casev ~ (s, \seq v) \\
%
% Wasm Context
  \text{Wasm Context}\quad& WC &\ni& wc &::=& ~ \casev ~ (\text{``Label"}, \seq v) ~ | ~ \casev ~ (\text{``Frame"}, \seq v) \\
%
% Wasm Instruction
  \text{Wasm Instruction}\quad& WI &\ni& wi &::=& ~ \casev ~ (s, \seq v) \\
%
% Store
  \text{Store}\quad& Store &\ni& sto &::=& ~ \seq{(s, v)} \\
%
% Continuation
  \text{Continuation}\quad & K &\ni& k &::=& ~ \mt &\quad\text{(Empty)} \\
    &&&& | & ~ \toplevelcall ~ (s, \seq v) &\quad\text{(Top-level Call)} \\
    &&&& | & ~ \call ~ (s, s, \seq v, k) &\quad\text{(Call)} \\
    &&&& | & ~ \exe ~ (\seq{wi}, k) &\quad\text{(Execute)} \\
    &&&& | & ~ \wasm ~ (\seq{wc}, k) &\quad\text{(Wasm)} \\
    &&&& | & ~ \algo ~ c &\quad\text{(Algorithm)} \\
    &&&& | & ~ \ret ~ (v, k) &\quad\text{(Return)} \\
%
% Context
  \text{Context}\quad& C &\ni& c &::=& ~ (\seq i, \mu, \seq{wc}, k) \\
%
% Environment
  \text{Environment}\quad& M &\ni& \mu &::=& ~ \seq{[s \mapsto v]} \\
\end{array}
\end{align*}

\newpage
\begin{align*}
\begin{array}{lcccrlr}
%
% Value
  \text{Value}\quad& V &\ni& v &::=& ~ \numv ~ n &\quad\text{(Number)} \\
    &&&& | & ~ \boolv ~ b &\quad\text{(Boolean)} \\
    &&&& | & ~ \fnamev ~ s &\quad\text{(Function Name)} \\
    &&&& | & ~ \listv ~ \seq v &\quad\text{(List)} \\
    &&&& | & ~ \strv ~ \seq{(s, v)} &\quad\text{(Record)} \\
    &&&& | & ~ \tupv ~ \seq v &\quad\text{(Tuple)} \\
    &&&& | & ~ \casev ~ (s, \seq v) &\quad\text{(Tagged Tuple)} \\
    &&&& | & ~ \trapv &\quad\text{(Trap)} \\
    &&&& | & ~ \storev &\quad\text{(Store)} \\
\end{array}
\end{align*}

% State
An AL state consists of a sequence of algorithms, a Wasm state, and a
continuation.
% Wasm State
A Wasm state consists of an interleaved stack, a Wasm instruction stack, and a
store.
% Wasm Entry
An interleaved stack contains two types of entries: a Wasm value and a Wasm
context.
% Wasm Value & Wasm Context & Wasm Instruction
Each of a Wasm value, a Wasm context, and a Wasm instruction is expressed as an
AL value $\casev$.
In particular, the tag of a Wasm context is either ``Label" or ``Frame".
% Store
A store is expressed as a sequence of pairs, each consisting of a field name
and an AL value.
% Continuation
A continuation is \mt{} for an empty continuation, \toplevelcall{} for
top-level function call, \call{} for a function call with let binding, \exe{}
for executing given Wasm instructions, \wasm{} for executing a Wasm instruction
in the instruction stack, \algo{} for interpreting AL instructions in an
algorithm, or \ret{} for returning from an AL function.
% Env
An environment is a finite mapping from variable names to AL values.
% Context
An AL context consists of a sequence of Wasm instruction, an environment, a
sequence of Wasm contexts, and a continuation.
% Value
An AL value is a number, boolean, function name, list, record, tuple, tagged tuple,
trap, or store.

% Terminology
In this section, AL-specific terms (state, context, value, function, and
instruction) are referred to simply as state, context, value, function, and
instruction, while Wasm terms remain unchanged.

\newpage
%% Semantics
\section{Semantics of AL}
\label{semantics}

%% Continuation

\begin{gather*}
\boxed{\leadsto \subseteq \Sigma \times \Sigma} \\
%
% TopLevelCall
\newline \\
  \\
  \hline
  (\seq{a}, w, \toplevelcall ~ (s, \seq v)) \leadsto (\seq{a}, w, \createalgo(\seq a, s, \seq v, \mt)) \\
%
% Call
\newline \\
  \\
  \hline
  (\seq{a}, w, \call ~ (s_{var}, s_{name}, \seq v, k)) \leadsto
  (\seq{a}, w, \createalgo(\seq a, s_{name}, \seq v, \call ~ (s_{var}, s_{name}, \seq v, k))) \\
%
% Execute-empty
\newline \\
  \\
  \hline
  (\seq{a}, w, \exe ~ (\epsilon, k)) \leadsto (\seq{a}, w, k) \\
%
% Execute-instr
\newline \\
  \casev ~ (s, \seq v) = wi \\
  \hline
  (\seq{a}, w, \exe ~ (wi ~ \seq{wi}, k)) \leadsto
  (\seq{a}, w, \createalgo(\seq a, s, \seq v, \exe ~ (\seq{wi}, k))) \\
%
% Wasm-empty
\newline \\
  \\
  \hline
  (\seq{a}, w, \wasm ~ (\epsilon, k)) \leadsto (\seq{a}, w, k) \\
%
% Wasm-instr
\newline \\
  (w', \casev ~ (s, \seq v)) = \popwasminstr(w) \\
  \hline
  (\seq{a}, w, \wasm ~ (wc^+, k))
  \leadsto
  (\seq{a}, w', \createalgo(\seq a, s, \seq v, \wasm ~ (wc^+, k))) \\
%
% Al-empty
\newline \\
  \\
  \hline
  (\seq{a}, w, \algo ~ (\epsilon, \mu, \seq{wc}, k)) \leadsto (\seq{a}, w, k) \\
%
% Al-instr
\newline \\
  \seq{a}, w, (\seq{i_1}, \mu, \seq{wc}, k) \vdash i_0 \Rightarrow (w', k') \\
  \hline
  (\seq{a}, w, \algo ~ (i_0 ~ \seq{i_1}, \mu, \seq{wc}, k)) \leadsto (\seq{a}, w', k') \\
%
% Return
\newline \\
  \mu = \getenv(c) \qquad c' = \setenv(c, \mu[s_{var} \mapsto v]) \\
  \hline
  (\seq{a}, w, \ret ~ (v, \call ~ (s_{var}, s_{name}, \seq v, \algo ~ c)) \leadsto
  (\seq a, w, \algo ~ c') \\
\end{gather*}

The semantics of AL is defined by a state transition system.
% TopLevelCall
Given algorithms $\seq a$ generated from the DSL and a Wasm state $w$, a
top-level function $s$ can be invoked with arguments $\seq v$ to perform
either module instantiation or function invocation:
$(\seq a, w, \toplevelcall ~ (s, \seq v))$.
$\toplevelcall ~ (s, \seq v)${} transitions to \algo{} with the function name
$s$, the arguments $\seq v$, and \mt{}.
% Empty
\mt{} is an empty continuation, representing the end of the transition, so no
further transition occurs.
% Call
Similarly to the top-level call, $\call~ (s_{var}, s_{name}, \seq v, k)${}
transitionsto \algo{} with the function name $s_{name}$ and the arguments $\seq
v$, but with the current continuation nested within it.
% Execute
$\exe ~ (\seq{wi}, k)${} executes each Wasm instruction in the sequence
$\seq{wi}$ one by one, each containing a relation name and arguments.
The continuation transitions to \algo{} with the relation name, the arguments,
and the current continuation, with the Wasm instruction being popped.
% Wasm
In $\wasm ~ (\seq{wc}, k)${}, a Wasm instruction is popped from
the Wasm instruction stack and executed until the Wasm context sequence
$\seq{wc}$ is exhausted.
This execution transitions the continuation to \algo{}, with the relation name
and the arguments of the Wasm instruction, and the current continuation nested
within it.
% Algo
In $\algo ~ (\seq i, \mu, \seq{wc}, k)${}, the body instructions $\seq i$
execute sequentially, allowing transitions to \call{}, \exe{}, \wasm{}, \ret{},
or \algo{}.
% Return
The execution of a function call concludes with a return instruction, which
changes the continuation to $\ret ~ v${}.
If the function call originates from \call{}, the return value $v$ is assigned
to the variable $s_{var}$.
If it originates from \toplevelcall{}, the inner continuation $k$ is always
\mt, so the entire execution concludes with the return value $v$:
$
(\seq a, w, \toplevelcall ~ (s, \seq v))
\leadsto^*
(\seq a, w', \ret ~ (v, \mt))
$.

%% Instruction

\begin{gather*}
  \boxed{\seq{a}, ~ w, ~ c \vdash i \Rightarrow w, ~ k} \\
%
% If-true
\newline \\
  w, \getenv(c) \vdash e \Rightarrow v \qquad
  \istrue(v) \\
  \hline
  \seq{a}, w, c \vdash \ifi ~ (e, \seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_1}))) \\
%
% If-false
\newline \\
  w, \getenv(k) \vdash e \Rightarrow v \qquad
  \neg \istrue(v) \\
  \hline
  \seq{a}, w, c \vdash \ifi ~ (e, \seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_2}))) \\
%
% Either-1
\newline \\
  \\
  \hline
  \seq{a}, w, c \vdash \eitheri ~ (\seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_1}))) \\
%
% Either-2
\newline \\
  \\
  \hline
  \seq{a}, w, c \vdash \eitheri ~ (\seq{i_1}, \seq{i_2}) \Rightarrow
  (w, \algo ~ (\prependinstr(c, \seq{i_2}))) \\
%
% Enter
\newline \\
  (\seq{we}, \seq{wi}, sto) = w \qquad
  (i^*, \mu, wc_1 ~ ... ~ wc_n, k) = c \\
  w, \mu \vdash e_1 \Rightarrow wc \qquad
  w, \mu \vdash e_2 \Rightarrow \listv ~ \seq{wi_1} \\
  wi = \getendinstr(wc) \qquad
  wi_1 = \getendinstr(wc_1) \quad ... \quad wi_n = \getendinstr(wc_n) \\
  wi_2^* = wi_1^* ~ wi ~ wi_1 ~ ... ~ wi_n \\
  \hline
  \seq{a}, w, c \vdash \enteri ~ (e_1, e_2)
  \Rightarrow
  (
    (wc ~ \seq{we}, wi_2^* ~ \seq{wi}, sto),
    \wasm ~ (wc ~ wc_1 ~ ... ~ wc_n, \algo ~ (i^*, \mu, \epsilon, k))
  ) \\
%
% PushCtx
\newline \\
  w, \getenv(c) \vdash e \Rightarrow wc \\
  \hline
  \seq{a}, w, c \vdash \pushctxi ~ e
  \Rightarrow
  (\push(w, wc), \algo ~ (\addctx(c, wc))) \\
%
% Push
\newline \\
  w, \getenv(c) \vdash e \Rightarrow wv \\
  \hline
  \seq{a}, w, c \vdash \pushi ~ e \Rightarrow (\push(w, wv), \algo ~ c) \\
%
% PopCtx
\newline \\
  (\seq{we}, \seq{wi}, sto) = w \qquad
  (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \\
  \mu = \getenv(c) \cdot \assign(e, wc) \qquad
  \seq{wi'} = \exit(\seq{wi}) \qquad
  c' = \popwasmctx_{C}(c) \\
  \hline
  \seq{a}, w, c \vdash \popctxi ~ e
  \Rightarrow
  ((\seq{we'}, \seq{wi'}, sto), \algo ~ (\setenv(c', \mu))) \\
%
% Pop
\newline \\
  (w', wv) = \pop(w) \qquad
  \mu' = \getenv(c) \cdot \assign(e, wv) \\
  \hline
  \seq{a}, w, c \vdash \popi ~ e \Rightarrow (w', \algo ~ (\setenv(c, \mu'))) \\
%
% PopN
\newline \\
  \mu = \getenv(c) \qquad
  w, \mu \vdash e_2 \Rightarrow \numv ~ n \\
  (w', wv^n) = \popn(w, n) \qquad
  \mu' = \mu \cdot \assign(e_1, \listv ~ wv^n) \\
  \hline
  \seq{a}, w, c \vdash \popni ~ (e_1, e_2) \Rightarrow
  (w', \algo ~ (\setenv(c, \mu'))) \\
%
% PopAll
\newline \\
  (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \qquad
  \mu' = \getenv(c) \cdot \assign(e, \liste ~ \seq{wv}) \\
  \hline
  \seq{a}, (\seq{we}, \seq{wi}, sto), c \vdash \popalli ~ e
  \Rightarrow
  ((wc ~ \seq{we'}, \seq{wi}, sto), \algo ~ (\setenv(c, \mu'))) \\
%
% Let
\newline \\
  \mu = \getenv(c) \qquad
  w, \mu \vdash e_2 \Rightarrow v \qquad
  \mu' = \mu \cdot \assign(e_1, v) \\
  \hline
  \seq{a}, w, c \vdash \leti ~ (e_1, e_2)
  \Rightarrow
  (w, \setenv(c, \mu')) \\
%
% Trap
\newline \\
  \\
  \hline
  \seq{a}, w, c \vdash \trapi \Rightarrow (w, \ret ~ (\trapv, \mt)) \\
%
% Return-rule
\newline \\
  \\
  \hline
  \seq{a}, w, (\seq{i}, \mu, wc^*, k) \vdash \returnreli \Rightarrow (w, k) \\
%
% Return-func
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  \seq{a}, w, (\seq{i}, \mu, wc^*, k) \vdash \returnfuni ~ e \Rightarrow
  (w, \ret ~ (v, k)) \\
%
% Execute
\newline \\
  w, \getenv(c) \vdash e \Rightarrow \listv ~ wi^* \\
  \hline
  \seq{a}, w, c \vdash \executei ~ e \Rightarrow
  (w, \exe ~ (wi^*, \algo ~ c)) \\
%
% Call-fname
\newline \\
  \mu = \getenv(c) \qquad
  \fnamev ~ s_3 = \mu(s_2) \qquad
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  \seq{a}, w, c \vdash \calli ~ (s_1, s_2, e_1 ~ ... ~ e_n) \Rightarrow
  (w, \call ~ (s_1, s_3, v_1 ~ ... ~ v_n, \algo ~ c)) \\
%
% Call-fname
\newline \\
  \mu = \getenv(c) \qquad
  s_2 \not\in \domain(\mu) \qquad
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  \seq{a}, w, c \vdash \calli ~ (s_1, s_2, \seq e) \Rightarrow
  (w, \call ~ (s_1, s_2, v_1 ~ ... ~ v_n, \algo ~ c)) \\
%
% Replace-frame
\newline \\
\text{NOTE: $e_1$ is either store or current frame} \\
  \mu = \getenv(c) \qquad
  \casev ~ (\text{``Frame"}, v_1 ~ v_2) = \getcurframe(w) \qquad
  w, \mu \vdash e \Rightarrow v \\
  wc = \casev ~ (\text{``Frame"}, v_1 ~ \update(w, \mu, v_2, p^+, v)) \\
  \hline
  \seq{a}, w, c \vdash \replaceframei ~ (p^+, e) \Rightarrow
  (\setcurframe(w, wc), \algo ~ c) \\
%
% Replace-store
\newline \\
  \mu = \getenv(c) \qquad
  w, \mu \vdash e \Rightarrow v \\
  \strv ~ sto = \update(w, \mu, \strv ~ (\getstore(w)), p^+, v) \\
  \hline
  \seq{a}, w, c \vdash \replacestorei ~ (p^+, e)
  \Rightarrow (\setstore(w, sto), c) \\
\end{gather*}

Given algorithms $\seq a$, a Wasm state $w$, and a context $c$, executing an
instruction $i$ produces an updated Wasm state and a continuation.
% If
$\ifi ~ (e, i_1^*, i_2^*)${} evaluates $e$ and branches to $\seq i_1$ if the
result is true, or $\seq i_2$ otherwise.
% Either
$\eitheri ~ (i_1^*, i_2^*)${} nondeterministically selects either $i_1^*$ or
$i_2^*$.
% Enter
$\enteri ~ (e_1, e_2)${} evaluates $e_1$, pushing the resulting Wasm context
$wc$ to the interleaved stack.
It then evaluates $e_2$ to obtain the Wasm instruction sequence $wi_1^*$.
Then, the end instructions of the Wasm context $wc$ and the Wasm contexts
$wc_1 ~ ... ~ wc_n$ in the context $c$ are appended to $wi_1^*$ and pushed to
the Wasm instruction stack.
Finally, it changes the continuation to
$\wasm ~ (wc ~ wc_1 ~ ... ~ wc_n, \algo ~ (i^*, \mu, \epsilon, k))${}.
% PushCtx
$\pushctxi ~ e${} evaluates $e$, pushing the resulting Wasm context $wc$ to the
interleaved stack and the context $c$.
% Push
$\pushi ~ e${} evaluates $e$, pushing the resulting value $wv$ to the
interleaved stack.
% PopCtx
$\popctxi ~ e${} pops all Wasm values up to and including a Wasm context $wc$
from the interleaved stack and all Wasm instructions up to and including an end
instruction from the Wasm instruction stack.
It also pops a Wasm context from the context $c$.
Then, it assign the Wasm context $wc$ to $e$.
% Pop
$\popi ~ e${} pops a Wasm value $wv$ from the interleaved stack and assign the
value to $e$.
% PopN
$\popni ~ (e_1, e_2)${} evaluates $e_2$ to obtain a number $n$, pops $n$ Wasm
values from the interleaved stack, and assigns the values to $e_1$.
% PopAll
$\popalli ~ e${} pops all the Wasm values up to, but not including, a
Wasm context from the interleaved stack and assign the values to $e$.
% Let
$\leti ~ (e_1, e_2)${} evaluates $e_2$ and assigns the resulting value to
$e_1$.
% Trap
\trapi{} results in $\ret ~ (\trapv{}, \mt{})$ to terminate the execution.
% ReturnRel
\returnreli{} results in the continuation $k$ in the context $c$, indicating a
return from a relation algorithm.
% ReturnFunc
$\returnfuni ~ e${} evaluates $e$ to $v$, resulting in $\ret ~ (v, k)${} where
$k$ is the continuation in the context $c$.
% Execute
\executei{} evaluates $e$ to obtain a Wasm instruction sequence $wi^*$ and
result in $\exe ~ (wi^*, \algo ~ c)${}.
% Call
$\calli ~ (s_1, s_2, e_1 ~ ... ~ e_n)${} checks if $s_2$ is bound in the
environment.
If it is, it retrieves the function name $s_3$ from the environment $\mu$;
otherwise, it uses $s_2$ as the function name.
It evaluates $\seq e$ to $v^*$, resulting in $\call ~ (s_1, s_{name}, v_1 ~ ...
~ v_n)${} where $s_{name}$ is the function name.
% ReplaceFrame
\replaceframei{} retrieves the current frame $\casev ~ (\text{``Frame"}, v_1
v_2)${} in the Wasm context $w$.
It then evaluates $e$ to $v$, updating the value at location $v_2$ along paths
$p^+$ to $v$.
% ReplaceFrame
\replacestorei{} evaluates $e$ to $v$, updating the value at location $\strv ~
sto$ along paths $p^+$ to v.

%% Expression

\begin{gather*}
  \boxed{w, \mu \vdash e \Rightarrow v} \\
%
% Var
\newline \\
  v = \mu(s) \\
  \hline
  w, \mu \vdash \vare ~ s \Rightarrow v \\
%
% Num
\newline \\
  \hline
  w, \mu \vdash \nume ~ n \Rightarrow \numv ~ n \\
%
% Bool
\newline \\
  \hline
  w, \mu \vdash \boole ~ b \Rightarrow \boolv ~ b \\
%
% Fname
\newline \\
  \hline
  w, \mu \vdash \fnamee ~ s \Rightarrow \fnamev ~ s \\
%
% List
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \liste ~ e_1 ~ ... ~ e_n \Rightarrow \listv ~ v_1 ~ ... ~ v_n \\
%
% Str
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \stre ~ (s_1, e_1) ~ ... ~ (s_n, e_n) \Rightarrow
  \strv ~ (s_1, v_1) ~ ... ~ (s_n, v_n) \\
%
% Tup
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \tupe ~ e_1 ~ ... ~ e_n \Rightarrow \tupv ~ v_1 ~ ... ~ v_n \\
%
% Case
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \quad ... \quad
  w, \mu \vdash e_n \Rightarrow v_n \\
  \hline
  w, \mu \vdash \casee ~ (s, e_1 ~ ... ~ e_n) \Rightarrow \casev ~ (s, v_1 ~ ... ~ v_n) \\
%
% Un
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  w, \mu \vdash \une ~ (unop, e) \Rightarrow \unop(unop, v) \\
%
% Bin
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad w, \mu \vdash e_2 \Rightarrow v_2 \\
  \hline
  w, \mu \vdash \bine ~ (binop, e_1, e_2) \Rightarrow \binop(binop, v_1, v_2) \\
%
% Acc
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  w, \mu \vdash \acce ~ (e, p) \Rightarrow \access(w, \mu, v, p) \\
%
% Upd
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad w, \mu \vdash e_2 \Rightarrow v_2 \\
  \hline
  w, \mu \vdash \upde ~ (e_1, p^+, e_2) \Rightarrow \update(w, \mu, v_1, p^+, v_2) \\
%
% Cat
\newline \\
   w, \mu \vdash e_1 \Rightarrow \listv ~ \seq{v_1} \qquad
   w, \mu \vdash e_2 \Rightarrow \listv ~ \seq{v_2} \\
  \hline
  w, \mu \vdash \cate ~ (e_1, e_2) \Rightarrow \listv ~ (\seq{v_1} ~ \seq{v_1}) \\
%
% Comp
\newline \\
   w, \mu \vdash e_1 \Rightarrow \strv ~ \seq{(s_1, v_1)} \qquad
   w, \mu \vdash e_2 \Rightarrow \strv ~ \seq{(s_2, v_2)} \\
  \hline
  w, \mu \vdash \compe ~ (e_1, e_2) \Rightarrow \strv ~ (\seq{(s_1, v_1)} ~ \seq{(s_2, v_2)}) \\
%
% Mem
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad
  w, \mu \vdash e_2 \Rightarrow \listv ~ \seq{v_2} \\
  \hline
  w, \mu \vdash \meme ~ (e_1, e_2) \Rightarrow \boolv ~ (v_1 \in \seq{v_2}) \\
%
% Choose
\newline \\
  w, \mu \vdash e \Rightarrow \listv ~ \seq{v} \qquad
  v \in \seq{v} \\
  \hline
  w, \mu \vdash \choosee ~ e \Rightarrow v \\
%
% Len
\newline \\
  w, \mu \vdash e \Rightarrow \listv ~ \seq{v} \\
  \hline
  w, \mu \vdash \lene ~ e \Rightarrow \numv ~ |\seq v| \\
%
% IsCaseOf
\newline \\
  w, \mu \vdash e \Rightarrow \casev ~ (s', \seq{v}) \\
  \hline
  w, \mu \vdash \iscaseofe ~ (e, s) \Rightarrow \boolv ~ (s = s') \\
%
% GetCurCtx
\newline \\
  wc = \getcurctx(w) \\
  \hline
  w, \mu \vdash \getcurctxe \Rightarrow wc \\
%
% CtxKind
\newline \\
  \casev (s', \seq v) = \getcurctx(w) \\
  \hline
  w, \mu \vdash \ctxkinde ~ s \Rightarrow \boolv ~ (s = s') \\
%
% Iter-
\newline \\
  \,~\mu(s_1) = \listv ~ (v_{11} ~ ... ~ v_{1n}) \quad ... \quad
  \mu(s_m) = \listv ~ (v_{m1} ~ ... ~ v_{mn}) \\
  \mu_1 = \mu[s_1 \mapsto v_{11}, ~ ... ~ s_m \mapsto v_{m1}] \quad ... \quad
  \mu_n = \mu[s_1 \mapsto v_{1n}, ~ ... ~ s_m \mapsto v_{mn}] \\
  w, \mu_1 \vdash e \Rightarrow v_1 \quad ... \quad
  w, \mu_n \vdash e \Rightarrow v_n \\
  \hline
  w, \mu \vdash \itere ~ (e, \listiter, s_1 ~ ... ~ s_m) \Rightarrow
  \listv ~ v_1 ~ ... ~ v_n \\
%
% Iter-n
\newline \\
  w, \mu \vdash e_2 \Rightarrow \numv ~ n \\
  \,~\mu(s_1) = \listv ~ (v_{11} ~ ... ~ v_{1n}) \quad ... \quad
  \mu(s_m) = \listv ~ (v_{m1} ~ ... ~ v_{mn}) \\
  \mu_1 = \mu[s_1 \mapsto v_{11}, ~ ... ~ s_m \mapsto v_{m1}] \quad ... \quad
  \mu_n = \mu[s_1 \mapsto v_{1n}, ~ ... ~ s_m \mapsto v_{mn}] \\
  w, \mu_1 \vdash e \Rightarrow v_1 \quad ... \quad
  w, \mu_n \vdash e \Rightarrow v_n \\
  \hline
  w, \mu \vdash \itere ~ (e_1, \listniter ~ e_2, s_1 ~ ... s_m) \Rightarrow
  \listv ~ v_1 ~ ... ~ v_n \\
%
% Iter-idx
\newline \\
  w, \mu \vdash e_2 \Rightarrow \numv ~ n \\
  \mu_1 = \mu[s_1 \mapsto \numv ~ 0, ~ ... ~ s_m \mapsto \numv ~ 0] \quad ... \quad
  \mu_n = \mu[s_1 \mapsto \numv ~ n, ~ ... ~ s_m \mapsto \numv ~ n] \\
  w, \mu_1 \vdash e \Rightarrow v_1 \quad ... \quad
  w, \mu_n \vdash e \Rightarrow v_n \\
  \hline
  w, \mu \vdash \itere ~ (e_1, \listidxiter ~ e_2, s_1 ~ ... ~ s_m) \Rightarrow
  \listv ~ v_1 ~ ... ~ v_n \\
%
% Match
\newline \\
  w, \mu \vdash e_1 \Rightarrow v_1 \qquad
  w, \mu \vdash e_2 \Rightarrow v_2 \\
  \hline
  w, \mu \vdash \matche ~ (e_1, e_2) \Rightarrow \boolv ~ (\match(v_1, v_2)) \\
%
% HasType
\newline \\
  w, \mu \vdash e \Rightarrow v \\
  \hline
  w, \mu \vdash \hastypee ~ (e, s) \Rightarrow \boolv ~ (\hastype(v, s)) \\
\end{gather*}

Given a Wasm state $w$ and an environment $\mu$, an expression $e$ evaluates
to a value.
% Var
$\vare ~ s${} looks up the environment $\mu$ and gets the value.
% Num Bool Fname
$\nume ~ n${}, $\boole ~ b${}, and $\fnamee ~ s${} directly reduce to $\numv ~
n${}, $\boolv ~ b${}, and $\fnamev ~ s${}, respectively.
% Tuple Case List Struct
$\liste ~ e_1 ~ ... ~ e_n${}, $\stre ~ (s_1, e_1) ~ ... ~ (s_n, e_n)${}, $\tupe
~ e_1 ~ ... ~ e_n${}, and $\casee ~ (s, e_1 ~ ... ~ e_n)${} evaluate their
sub-expressions and reduce to $\listv ~ v_1 ~ ... ~ v_n${}, $\strv ~ (s_1, v_1)
~ ... ~ (s_n, v_n)${}, $\tupv ~ v_1 ~ ... ~ v_n${}, and $\casev ~ (s, v_1 ~ ...
~ v_n)${}, respectively.
% Un Bin
$\une ~ (unop, e)${} and $\bine ~ (binop, e_1, e_2)${} evaluate their
sub-expressions and perform \unop{} and \binop{}, respectively.
% Acc
$\acce ~ (e, p)${} evaluates $e$ and accesses the resulting value $v$ using a
path $p$.
% Upd
$\upde ~ (e_1, p^+, e_2)${} evaluates $e_1$ and $e_2$ to $v_1$ and $v_2$,
updating the value at location $v_1$ along paths $p^+$ to the value $v_2$.
% Cat Comp
$\cate ~ (e_1, e_2)${} and $\compe ~ (e_1, e_2)${} each evaluate two
sub-expressions $e_1$ and $e_2$, concatenating the resulting lists and
resulting records, respectively.
% Mem
$\meme ~ (e_1, e_2)${} evaluates $e_1$ and $e_2$ to $v_1$ and $\listv ~ \seq
v_2$, and checks whether $v_1$ is a member of $\seq v_2$.
% Choose Len
$\choosee ~ e$ and $\lene ~ e${} each evaluate $e$ to the list value $\listv ~
\seq v$, choosing an element from the list and calculating the length of the
list, respectively.
% IsCaseOf
$\iscaseofe ~ (e, s)${} evaluates $e$ to the tagged tuple $\casev ~ (s', \seq
v)$ and checks whether the tag $s'$ matches $s$.
% GetCurCtx
\getcurctxe{} retrieves the current Wasm context from the interleaved stack.
% CtxKind
$\ctxkinde ~ s${} retrieves the current Wasm context $\casev ~ (s', \seq v)$
and checks whether the kind $s'$ matches $s$.
% Iter
$\itere ~ (e, iter, s^*)${} behaves differently depending on $iter$.
\listiter{} and $\listniter ~ e_2${} are essentially equivalent.
They assume that all the values mapped to the variables $\seq s$ by the
environment $\mu${} are list values with the same length $n$.
$\listniter ~ e_2${} checks whether the number $n$ matches the result of
evaluating $e_2$, while \listiter{} allows any number.
It generates a sequence of environments $\mu_1 ~ ... ~ \mu_n$ where the k-th
environment maps each variable to the k-th elements of each list.
Each environment is used to evaluate $e_1$, producing a sequence of values $v_1
~ ... ~ v_n$.
Finally, it results in $\listv ~ v_1 ~ ... ~ v_n$.
$\listidxiter ~ e_2${} behaves similarly except that the variables $\seq s$ are
unbound.
It evaluates $e_2$ to obtain $n$ and then behaves like $\listniter ~ e_2${} as
if a list of values [0 ... n-1] were bound to the variables $s^*$.
% IsValid Match HasType
$\matche ~ (e_1, e_2)${} and $\hastypee ~ (e, s)${} evaluate their
sub-expressions and perform \match{} and \hastype{}, respectively.
Note that \match{} and \hastype{} represent the match relation and the type
relation in the Wasm static semantics.
Their formalization is omitted as they are highly specific to Wasm and
excessively verbose.

%% Helper functions

\begin{align*}
%
% get_algo_name
\newline \\
  &\getalgoname(\rel ~ (s, \seq e, \seq i)) = s \\
  &\getalgoname(\fun ~ (s, \seq e, \seq i)) = s \\
%
% lookup
\newline \\
  &\lookup(\epsilon, s) = \epsilon \\
  &\lookup(a ~ \seq{a}, s) =
    \begin{cases}
      a &\quad\quad \premise{if}\quad ~ s = \getalgoname(a) \\
      \lookup(\seq{a}, s) &\quad\quad \premise{otherwise}
    \end{cases}
  \\
%
% assign
\newline \\
  &\assign(\vare ~ s, v) = [s \mapsto v] \\
  &\assign(\itere ~ (e, iter, s^*), v) = \assign(e, v) \\
  &\assign(\tupe ~ e_1 ~ ... ~ e_n, \tupv ~ v_1 ~ ... ~ v_n) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(\liste ~ e_1 ~ ... ~ e_n, \listv ~ v_1 ~ ... ~ v_n) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(\casee ~ (s, e_1 ~ ... ~ e_n), \casev ~ (s, v_1 ~ ... ~ v_n)) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(
    w,
    \mu,
    \stre ~ (s_1, e_1) ~ ... ~ (s_2, e_n),
    \strv ~ (s_1, v_1) ~ ... ~ (s_n, v_n)
  ) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_n, v_n) \\
  &\assign(\cate ~ (\liste ~ e_1 ~ ... ~ e_m, e), \listv v_1 ~ ... ~ v_n) =
    \assign(e_1, v_1) \cdot ~ ... ~ \cdot \assign(e_m, v_m) \cdot
    \assign(e, \listv ~ v_{m+1} ~ ... ~ v_n) \\
  &\assign(\cate ~ (e, \liste ~ e_1 ~ ... ~ e_m), \listv v_1 ~ ... ~ v_n) =
    \assign(e, \listv ~ v_1 ~ ... ~ v_{n-m}) \cdot
    \assign(e_1, v_{n-m+1}) \cdot ~ ... ~ \cdot \assign(e_m, v_n) \\
%
% create_algo
\newline \\
  &\createalgo(\seq a, s, \seq v, k) =
  \algo(\seq i, \mu, \epsilon, k) \\
  &\qquad\qquad\qquad\qquad\premise{if}\quad
  (\rel ~ (s, \seq e, \seq i) \quad\lor\quad \fun ~ (s, \seq e, \seq i)) = \lookup(\seq a, s)
  \quad\land\quad
  \mu = \assign(w, [\text{``s"} \mapsto \storev], \seq e, \seq v) \\
%
% split_ctx
\newline \\
  &\splitctx(\epsilon) = (\epsilon, \epsilon, \epsilon) \\
  &\splitctx(we ~ \seq{we}) =
    \begin{cases}
      (\epsilon, wc, \seq{we})
      &\quad\quad \premise{if}\quad
      we = wc \\
      (wv ~ \seq{wv}, wc, \seq{we'})
      &\quad\quad \premise{if}\quad
      we = wv \quad\land\quad
      (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \\
    \end{cases}
  \\
%
% pop-wasm-ctx_context
\newline \\
  &\popwasmctx_{C}((\seq i, \mu, epsilon, k)) = (\seq i, \mu, \epsilon, \popwasmctx_K(k)) \\
  &\popwasmctx_{C}((\seq i, \mu, wc ~ \seq{wc}, k)) = (\seq i, \mu, wc^*, k) \\
%
\newline \\
% pop-wasm-ctx_cont-mt
  &\popwasmctx_{K}(\mt) = \epsilon \\
% pop-wasm-ctx_cont-toplevelcall
  &\popwasmctx_{K}(\toplevelcall ~ (s, v^*)) = \epsilon \\
% pop-wasm-ctx_cont-call
  &\popwasmctx_{K}(\call ~ (s_1, s_2, \seq v, k)) =
    \call ~ (s_1, s_2, \seq v, (\popwasmctx_K(k)) \\
% pop-wasm-ctx_cont-execute
  &\popwasmctx_{K}(\exe ~ (v, k)) = \exe ~ (v, (\popwasmctx_K(k)) \\
% pop-wasm-ctx_cont-wasm_0
  &\popwasmctx_{K}(\wasm ~ (\epsilon, k)) = \wasm ~ (\epsilon, \popwasmctx_{K}(k)) \\
% pop-wasm-ctx_cont-wasm_nonzero
  &\popwasmctx_{K}(\wasm ~ (wc ~ \seq{wc}, k)) = \wasm ~ (\seq{wc}, k) \\
% pop-wasm-ctx_cont-al
  &\popwasmctx_{K}(\algo ~ c) = \algo ~ (\popwasmctx_C(c)) \\
% pop-wasm-ctx_cont-ret
  &\popwasmctx_{K}(\ret ~ (v, k)) = \ret ~ (v, \popwasmctx_K(k)) \\
%
% exit
\newline \\
  &\exit(wi ~ \seq{wi}) =
    \begin{cases}
      \seq{wi} &\quad\quad \premise{if}\quad ~ \isendinstr(wi) \\
      \exit(\seq{wi}) &\quad\quad \premise{otherwise} \\
    \end{cases}
  \\
%
% get_env
\newline \\
  &\getenv((\seq i, \mu, wc^*, k)) = \mu \\
%
% set_env
\newline \\
  &\setenv((\seq i, \mu, wc^*, k), \mu') = (\seq i, \mu', wc^*, k) \\
%
% prepend_instr
\newline \\
  &\prependinstr((\seq i, \mu, wc^*, k), \seq{i'}) = (\seq{i'} ~ \seq i, \mu, wc^*, k) \\
%
% add_ctx
\newline \\
  &\addctx((\seq i, \mu, wc^*, k), wc') = (\seq i, \mu, wc' ~ wc^*, k) \\
%
% get_store
\newline \\
  &\getstore((\seq{we}, \seq{wi}, sto)) = sto \\
%
% set_store
\newline \\
  &\setstore((\seq{we}, \seq{wi}, sto), sto') = (\seq{we}, \seq{wi}, sto') \\
%
% pop_wasm_instr
\newline \\
  &\popwasminstr((\seq{we}, wi ~ \seq{wi}, sto)) = ((\seq{we}, \seq{wi}, sto), wi) \\
%
% push
\newline \\
  &\push((\seq{we}, \seq{wi}, sto), we) = (we ~ \seq{we}, \seq{wi}, sto) \\
%
% pop
\newline \\
  &\pop((we ~ \seq{we}, \seq{wi}, sto)) = ((\seq{we}, \seq{wi}, sto), we) \\
%
% popn
\newline \\
  &\popn((we^n ~ \seq{we}, \seq{wi}, sto), n) = ((\seq{we}, \seq{wi}, sto), we^n) \\
%
% unop
\newline \\
  &\unop(\notop, v) =
    \begin{cases}
      \boolv ~ true &\quad\quad\premise{if}\quad ~ \istrue(v) \\
      \boolv ~ false &\quad\quad\premise{otherwise} \\
    \end{cases}
  \\
  &\unop(\minusop, \numv ~ n) = \numv ~ (-n) \\
%
% get_cur_ctx
\newline \\
  &\getcurctx((\seq{we}, \seq{wi}, sto)) =
  wc \quad\quad\premise{if}\quad ~ (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \\
%
% get_cur_frame-frame
\newline \\
  &\getcurframe((\seq{we}, \seq{wi}, sto)) =
    \begin{cases}
      wc &\quad\quad\premise{if}\quad
      (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad\land\quad \isframe(wc)\\
      \getcurframe(\seq{we'}) &\quad\quad\premise{if}\quad
      (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad\land\quad \neg \isframe(wc) \\
    \end{cases}
  \\
%
% set_cur_frame-frame
\newline \\
  &\setcurframe((\seq{we}, \seq{wi}, sto), wc_{frame})
  =
  \begin{cases}
    (\seq{wv} ~ wc_{frame} ~ \seq{we'}, \seq{wi}, sto) \\
    \qquad\qquad\qquad\quad\premise{if}\quad
    (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad \isframe(wc) \\
    (\seq{wv} ~ wc ~ \setcurframe(\seq{we'}, wc_{frame}), \seq{wi}, sto) \\
    \qquad\qquad\qquad\quad\premise{if}\quad
    (\seq{wv}, wc, \seq{we'}) = \splitctx(\seq{we}) \quad \neg \isframe(wc) \\
  \end{cases}
  \\
%
% access-idx
\newline \\
  &\access(w, \mu, \listv ~ \seq{v}, \idxp ~ e) = \seq{v}[n]
  \qquad\qquad\qquad~\,\quad\premise{if}\quad w, \mu \vdash e \Rightarrow \numv ~ n \\
% access-slice
  &\access(w, \mu, \listv ~ \seq{v}, \slicep ~ (e_1, e_2)) = \seq{v}[n_1: n_2]
  \quad\quad\premise{if}\quad
  w, \mu \vdash e_1 \Rightarrow \numv ~ n_1 \quad\land\quad
  w, \mu \vdash e_2 \Rightarrow \numv ~ n_2 \\
% access-dot
  &\access(w, \mu, \strv ~ ((s_0, v_0) ~ \seq{(s_1, v_1)}), \dotp ~ s) =
  \begin{cases}
    v_0 &\quad\quad\premise{if}\quad s_0 = s \\
    \access(w, \mu, \strv ~ \seq{(s_1, v_1)}, \dotp ~ s) &\quad\quad\premise{otherwise} \\
  \end{cases}
  \\
% access-store
  &\access(w, \mu, \storev, p) =\access(w, \mu, \strv ~ sto, p)
  \quad\quad\premise{if}\quad w = (\seq{we}, \seq{wi}, sto)\\
%
% update
\newline \\
  % idx
  &\update(w, \mu, \listv ~ \seq{v}, (\idxp ~ e) ~ \seq{p}, v) =
  \listv ~ (\updateidx(\seq v, n, \update(w, \mu, \seq{v}[n], \seq{p}, v))) \\
  &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\premise{if}\quad
  w, \mu \vdash e \Rightarrow \numv ~ n \\
  % slice
  &\update(w, \mu, \listv ~ \seq{v}, (\slicep ~ e) ~ \seq{p}, v) =
  \listv ~ (\updateslice(\seq v, n_1, n_2, \update(w, \mu, \seq{v_i}[n_1: n_2], \seq{p}, v))) \\
  &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\premise{if}\quad
  w, \mu \vdash e_1 \Rightarrow \numv ~ n_1 \quad\land\quad
  w, \mu \vdash e_2 \Rightarrow \numv ~ n_2 \\
  % dot
  &\update(w, \mu, \strv ~ \seq{(s, v)}, (\dotp ~ s') ~ \seq{p}, v)
  =
  \strv ~ (\updatedot(\seq{(s, v)}, s', \update(w, \mu, \access(v, \dotp ~ s'), \seq{p}, v))) \\
%
% update_idx
\newline \\
  &\updateidx(v_0 ~ \seq{v_1}, n, v_2) =
  \begin{cases}
    v_2 ~ \seq{v_1}
    &\quad\quad\premise{if}\quad n = 0 \\
    v_0 ~ \updateidx(\seq{v_1}, n-1, v_2)
    &\quad\quad\premise{if}\quad n > 0 \\
  \end{cases} \\
%
% update_slice
\newline \\
  &\updateslice(v_0^n ~ \seq{v_1}, 0, n, v_2^n) = v_2^n ~ \seq{v_1} \\
  &\updateslice(v_0 ~ \seq{v_1}, m, n, v_2^n) =  v_0 ~ \updateslice(\seq{v_1}, m-1, n, v_2)
  \quad\quad\premise{if}\quad m > 0 \\
%
% update_dot
\newline \\
  &\updatedot((s_0, v_0) ~ \seq{(s_1, v_1)}, s_2, v_2) =
  \begin{cases}
    (s_2, v_2) ~ \seq{(s_1, v_1)}
    &\quad\quad\premise{if}\quad s_0 = s_2 \\
    (s_0, v_0) ~ \updatedot(\seq{(s_1, v_1)}, s_2, v_2)
    &\quad\quad\premise{otherwise}
  \end{cases}
  \\
%
% get_end_algo
\newline \\
  &\getendinstr(wc) =
  \begin{cases}
    \casev ~ (\text{``Exiting"}, \epsilon) &\quad\quad\premise{if}\quad wc = \casev ~ (\text{``Label"}, \seq v)\\
    \casev ~ (\text{``Returning"}, \epsilon) &\quad\quad\premise{if}\quad wc = \casev ~ (\text{``Frame"}, \seq v)
  \end{cases} \\
%
% is_end_instr
\newline \\
  &\isendinstr(wi) =
  \begin{cases}
    true &\quad\quad\premise{if}\quad wi = \casev ~ (\text{``Exiting"}, \epsilon) \quad\lor\quad wi = \casev ~ (\text{``Returning"}, \epsilon)\\
    false &\quad\quad\premise{otherwise}
  \end{cases} \\
%
% is_true
\newline \\
  &\istrue(v) =
  \begin{cases}
  b &\quad\quad\premise{if}\quad v = \boolv ~ b \\
    \istrue(v_1) ~ \land ~ ... ~ \land ~ \istrue(v_n) &\quad\quad\premise{if}\quad
    v = \listv ~ v_1 ~ ... ~ v_n \\
  \end{cases} \\
%
% is_frame
\newline \\
  &\isframe(v) =
  \begin{cases}
    true
    &\quad\quad\premise{if}\quad v = \casev ~ (\text{"FRAME"}, \seq v) \\
    false &\quad\quad\premise{otherwise}\\
  \end{cases}
  \\
\end{align*}