### 2024-11-25 paper/eval.tex
Add eval

% !TEX root = main.tex

\chapter{Evaluation}
\label{ch:eval}
\noindent

We evaluate our approach with the following two research questions.
\begin{itemize}
  \item \textbf{RQ1. (Correctness)} Does AL ensures the correct specification?
    (\S 5.1)
  \item \textbf{RQ2. (Generality)} Does AL capture the fundamental operations of
    Wasm to describe Wasm semantics rather than being tailored to the current
    version of the specification? (\S 5.2)
\end{itemize}

\section{RQ1: Correctness}

We define the syntax and semantics of Wasm manually using the DSL.
The DSL is translated into AL, which serves as a Wasm interpreter that
faithfully follows the written specification.
To ensure correctness of the specification, we run this Wasm interpreter
against the execution tests from the official Wasm test suite.
The execution tests check whether the execution of a Wasm program produces
expected values or errors.
Tests related to inifinite loop detection are excluded, as the specification
does not define this behavior.
The results demonstrate that the interpreter successfully pass all 51,549 tests in 70
seconds on \red{(lambda 9)}.
Passing all the tests ensures the correctness of the specification.

% 7 + 3 + ??? + 44
\section{RQ2: Generality}
We evaluate AL's generality by applying SpecTec to eight proposals prepared for
inclusion in the next version of Wasm: typed function references
~\cite{proposal-fr}, garbage collection~\cite{propsal-gc}, tail calls
~\cite{proposal-tc}, multiple memories~\cite{proposal-mm}, extended constant
expressions~\cite{proposal-ec}, memory64~\cite{proposal-m64}, relaxed SIMD
~\cite{proposal-rs}, exception handling~\cite{proposal-eh}.
This proposals add or modify ??? Wasm instructions.
In particular, the garbage collection proposal adds substantial complications
to the Wasm type system, and the exception handling proposal introduces a new
context, \textit{handler}.
Notably, specifying all eight proposals in AL requires only trivial
adjustments, such as adding a handler case to the context.

We manually craft the specification including the proposal features using the
DSL, and similarly generate a Wasm interpreter using AL.
This interpreter is tested against the execution tests in the official test
suites and the proposals, excluding test related to the infinite loop
detection.
The interpreter successfully passed all ??? execution tests, in ??? seconds.
Passing all these tests demonstrates the generality of AL.