### 2024-11-26 paper/conclusion.tex
Write abstract and conclusion

% !TEX root = main.tex

\chapter{Conclusion}
\label{ch:conclusion}
\noindent

This thesis introduces SpecTec, a framework designed to mechanize the
specification of Wasm.
It offers a domain specific language that facilitates the declarative
definition of Wasm's syntax and semantics, closely resembling formal notation.
SpecTec incorporates type-checking within the DSL to prevent errors in
meta-level specifications and translates them into prose specifications.

To formally define the semantics of Wasm in the prose specification, we propose
a rigorous model of Wasm’s control flow, define the syntax and semantics of the
AL language for specifying the prose, and implement an AL interpreter.
By leveraging this interpreter, specifications written in AL are executable,
functioning as a Wasm interpreter.

Executing these specifications enables the direct evaluation of the Wasm
specification.
This framework bridges the gap between formal specification and executable
code, ensuring their correctness and fidelity.