### 2024-11-29 paper/intro.tex
Fix intro

% !TEX root = main.tex

\chapter{Introduction}
\label{ch:intro}
\noindent

% Brief overview
WebAssembly (Wasm)~\cite{wasm} is a low-level bytecode language that is safe,
fast, portable, and compact.
It is widely used as a compilation target for web applications.
Beyond web development, Wasm's advantages are also deployed in areas such as
cloud and edge computing~\cite{lucet, cloudflare}, IoT~\cite{wasm-iot}, and
blockchains~\cite{wasm-blockchain}.
% Risk of implementation divergence
There are numerous Wasm engines, with all major browsers implementing their own
multi-tiered versions~\cite{v8, spidermonkey, webkit}.
Additionally, specialized engines target specific domains, such as embedded
systems and edge computing.
However, ensuring portability across these diverse implementations introduces
the risk of divergence among them.

% Rigorous standardization -> formal notation & prose notation
To address this challenge, Wasm has been rigorously standardized by the
W3C~\cite{wasm-w3c}.
The Wasm specification is particularly rigorous, describing its semantics in
two complementary forms: \textit{formal notation} and \textit{prose notation}.
\cref{fig:testop} illustrates the specification document for $testop$
instruction, where the upper part contains the prose notation, and the lower
part contains the formal notation.
The formal notation employs mathematical rules to succinctly define the
semantics, supporting proofs such as type soundness.
Conversely, the prose notation uses pseudocode-like algorithms to explain the
semantics through step-by-step instructions.
Since most Wasm users and engine developers are less familiar with mathematical
formalism, they primarily rely on the prose notation.

\begin{figure}[t]
    \centerline{\includegraphics[width=15cm]{fig/testop}}
    \caption[\texttt{testop} instruction in official specification document]
      {\texttt{testop} instruction in official specification document}
      \label{fig:testop}
    \centerline{\includegraphics[width=15cm]{fig/spectec-testop}}
    \caption[\texttt{testop} instruction in SpecTec specification document]
      {\texttt{testop} instruction in SpecTec specification document}
      \label{fig:spectec-testop}
\end{figure}

% Challenging specification process
The demanding standardization process places a significant burden on
specification authors, particularly in maintaining consistency and correctness.
Crafting the specification document is labor-intensive, which becomes
increasingly difficult to scale as Wasm evolves.
This challenge is compounded by the dual requirement to maintain both a formal
specification, written in LaTeX, and a prose specification, authored in
reStructuredText.
The separate and user-unfriendly nature of these formats hinders collaborative
review and increases the risk of discrepancies and errors between the two
notations, further complicating the standardization process.

% SpecTec
To mitigate this problem, we propose SpecTec~\cite{spectec}, a framework for
mechanizing WebAssembly specification.
It provides domain specific language (DSL) that enables the declarative
definition of Wasm syntax and semantics, akin to the formal notation.
SpecTec performs type-checking on the DSL to prevent meta-level specification
errors and generates LaTeX for the formal notation.
Additionally, to capture the pseudocode-like structure of the prose notation,
SpecTec incorporates an imperative language, \textit{AL}, which stands for
Algorithmic Language.
SpecTec translates the DSL into AL and generate reStructuredText for the prose
notation.
\cref{fig:spectec-testop} is the specification document generated by the
SpecTec using the DSL and the AL.
The generated document closely resembles the official document except for some
minor notation changes and some missing hyperlinks.
One notable aspect of AL is that it is executable, meaning any specification
written in AL serves as a Wasm interpreter program.
By testing the Wasm interpreter program, we can check the correctness of the
specification.
This bridges the gap between formal specification and executable code.
Futhermore, SpecTec generates mechanized definitions for theorem provers.

% 느낌 대신 "논리"적인 흐름에 집중 (일관성과 일맥상통)
% Challenge
When desiging the SpecTec, one of the challenges in our approach is making AL
executable.
In this master thesis, we focus on this specific challenge.
Since the prose specification is written in informal pseudocode, its
interpretation is not straightforward.
The absence of a formalized definition or formal semantics for the prose
specification posed significant difficulty in realizing executable AL.
Since executing the specification against tests is the primary method for
assessing its correctness, this limitation hindered the development of both the
generated interpreter and the specification itself.

% Solution
To address this, we present a formal model for how the prose specification
describes Wasm semantics.
Based on these insight, we formally define AL's syntax and semantics to
accurately capture Wasm semantics.
We then implement an AL interpreter according to the formal definition and
evaluate its correctness.
The interpreter successfully passes all tests in the Wasm official test suite,
as well as tests from proposals.

% Contribution
The contributions of this paper are as follows:
\begin{itemize}
  \item We propose a formal model for the prose specification.
    (\cref{sec:model})
  \item We define syntax and semantics of AL (\cref{sec:definition})
  \item We implement an AL interpreter for executable specification
    (\cref{sec:exec-spec})
\end{itemize}